### 2020-09-16 exploratory/2020-09-09-genome-sens-spec.Rmd
Combine the count_tibble files and update genome-level exploratory analysis, closes #23

---
title: "Analyzing the sensitivity and specificity of ASVs for discriminating between genomes"
author: "Pat Schloss"
date: "9/9/2020"
output:
  github_document:
    html_preview: false
editor_options: 
  chunk_output_type: console
---

```{r, message=FALSE}
library(tidyverse)
library(here)
```

### Need to determine the numbe of *rrn* operons across genomes

Our analysis will use full length sequences

```{r}
count_tibble <- read_tsv(here("data/processed/rrnDB.count_tibble"),
												 col_types = "cccd")
```

We want to count and plot the number of copies per genome

```{r n_rrn}
count_tibble %>%
	filter(region == "v19") %>%
	group_by(genome) %>%
	summarize(n_rrn = sum(count), .groups="drop") %>%
	ggplot(aes(x=n_rrn)) + geom_histogram(binwidth=1)

count_tibble %>%
	filter(region == "v19") %>%
	group_by(genome) %>%
	summarize(n_rrn = sum(count)) %>%
	count(n_rrn) %>%
	mutate(fraction = n / sum(n))
```

We see that most genomes actually have more than one copy of the *rrn* operon. I wonder whether those different copies are the same sequence / ASV...

### Determine number of ASVs per genome

Considering most genomes have multiple copes of the *rrn* operon, we need to know whether they all have the same ASV. Otherwise we run the risk of splitting a single genome into multiple ASVs.

```{r}
count_tibble %>% 
	group_by(region, genome) %>%
	summarize(n_asv = n(), n_rrn = sum(count), .groups="drop") %>%
	group_by(region, n_rrn) %>%
	summarize(med_n_asv = median(n_asv),
						mean_n_asv = mean(n_asv),
						lq_n_asv = quantile(n_asv, prob=0.25),
						uq_n_asv = quantile(n_asv, prob=0.75)) %>%
	filter(n_rrn == 7)

count_tibble %>% 
	group_by(region, genome) %>%
	summarize(n_asv = n(), n_rrn = sum(count), .groups="drop") %>%
	ggplot(aes(x=n_rrn, y=n_asv, color=region)) + geom_smooth(method="lm")
```

Surprisingly (or not!) the number of ASVs increases at a rate of about 2 ASVs per 3 copies of *rrn* operon in the genome. The sub regions of the 16S rRNA region have few ASVs per *rrn* operon.

### Determine whether an ASV is unique to genomes they're found in

Instead of looking at the number of ASVs per genome, we want to see the number of genomes per ASV.

```{r}
count_tibble %>%
	group_by(region, asv) %>%
	summarize(n_genomes = n()) %>%
	count(n_genomes) %>%
	mutate(fraction = n/sum(n)) %>%
	filter(n_genomes == 1)
```

We see that will full length sequences, that 82% of the ASVs were unique to a genome. For the subregions, about 76% of the ASVs were unique to a genome.

### To be determined...
* Can correct for over representation? 
* Consider analysis at species, genus, family, etc. levels
* Consider looking at more broad definition of an ASV