### 2020-09-29 exploratory/2020-09-29-taxa-representation.Rmd
Determine number of genomes and taxa represented at each taxonomic level, closes #27

---
title: "Analyzing the distribution of genomes across different taxonomic ranks"
author: "Pat Schloss"
date: "9/29/2020"
output:
  github_document:
    html_preview: false
editor_options:
  chunk_output_type: console
---

```{r, message=FALSE}
library(tidyverse)
library(here)

metadata <- read_tsv(here("data/references/genome_id_taxonomy.tsv"),
										 col_types = cols(.default = col_character())) %>%
	mutate(strain = if_else(scientific_name == species,
													NA_character_,
													scientific_name)) %>%
	select(-scientific_name) %>% 
	pivot_longer(-genome_id, names_to="rank", values_to="taxon") %>%
	drop_na(taxon) %>%
	mutate(rank = factor(rank,
											 levels=c("kingdom", "phylum", "class", "order",
											 				 "family", "genus", "species", "strain")))

asv <- read_tsv(here("data/processed/rrnDB.count_tibble"),
								col_types = cols(.default = col_character(),
																 count = col_integer()))

metadata_asv <- inner_join(metadata, asv, by=c("genome_id" = "genome"))
```

### Find the number of taxa within each taxonomic rank

```{r}
n_taxa_per_rank <- metadata_asv %>%
	filter(region == "v19") %>% 
	group_by(rank, taxon) %>%
	summarize(N = n_distinct(genome_id)) %>%
	summarize(n_1 = n_distinct(taxon),
						n_2 = sum(N >= 2),
						n_3 = sum(N >= 3),
						n_4 = sum(N >= 4),
						n_5 = sum(N >= 5),
						n_10 = sum(N >= 10)
						) %>%
	pivot_longer(-rank, names_to = "n_genomes", values_to="n_taxa") %>%
	mutate(n_genomes = factor(n_genomes,
														levels=c("n_1", "n_2", "n_3", "n_4", "n_5", "n_10")))

n_taxa_per_rank %>% 
	ggplot(aes(x=rank, y=n_taxa, group=n_genomes, color=n_genomes)) +
		geom_line() + 
		theme_classic()
```

* Even if we require that ever taxonomic group had at least 5 genomes, there would be a few hundred species represented
* Not sure I trust the strain level data as being complete. There maybe taxa in the strain rank that are in the species but the people that deposited the data didn't indicate the correct strain name