### 2020-10-07 exploratory/2020-10-05-rrn-copy-number.Rmd
Add line segment to cloud of points to indicate average, closes #29

---
title: "Quantifying the number of *rrn* operons across taxonomic ranks"
author: "Pat Schloss"
date: "10/05/2020"
output:
  github_document:
    html_preview: false
---

```{r, message=FALSE}
library(tidyverse)
library(here)

metadata <- read_tsv(here("data/references/genome_id_taxonomy.tsv"),
										 col_types = cols(.default = col_character())) %>%
	mutate(strain = if_else(scientific_name == species,
													NA_character_,
													scientific_name)) %>%
	select(-scientific_name)

asv <- read_tsv(here("data/processed/rrnDB.count_tibble"),
								col_types = cols(.default = col_character(),
																 count = col_integer()))

metadata_asv <- inner_join(metadata, asv, by=c("genome_id" = "genome"))
```

### Plot the number of *rrn* copies per taxonomic rank

Our analysis will use full length sequences. We want to count and plot the number of copies per *taxonomic rank*. Before calculating the averages of each taxonomic group, we should first get an average number of copies for each species. This will allow us to control for uneven number of genomes in each species.

```{r}
rank_taxon_rrns <- metadata_asv %>%
	filter(region == "v19") %>%
	group_by(kingdom, phylum, class, order, family, genus, species, genome_id) %>%
	summarize(n_rrns = sum(count), .groups="drop") %>% 
	group_by(kingdom, phylum, class, order, family, genus, species) %>%
	summarize(mean_rrns = mean(n_rrns), .groups="drop") %>% 
	pivot_longer(-mean_rrns, names_to="rank", values_to="taxon") %>%
	drop_na(taxon) %>%
	mutate(rank = factor(rank,
											 levels=c("kingdom", "phylum", "class", "order",
											 				 "family", "genus", "species", "strain"))) %>%
	group_by(rank, taxon) %>%
	summarize(mean_rrns = mean(mean_rrns), .groups="drop")

mean_of_means <- rank_taxon_rrns %>%
	group_by(rank) %>%
	summarize(mean_mean_rrns = mean(mean_rrns), .groups="drop")

jitter_width <- 0.3
n_ranks <- nrow(mean_of_means)

rank_taxon_rrns %>%
	ggplot(aes(x=rank, y=mean_rrns)) + 
	geom_jitter(width=jitter_width, color="gray") +
	geom_segment(data=mean_of_means,
							 aes(x=1:n_ranks-jitter_width, xend=1:n_ranks+jitter_width,
							 		y=mean_mean_rrns, yend=mean_mean_rrns),
						 color="red", group=1, size=2, lineend="round",
						 inherit.aes=FALSE) + 
	theme_classic() +
	labs(x=NULL, 
			 y="Mean number of rrn copies per genome",
			 title="There is wide variation at the species level in the number of rrn copies",
			 subtitle="Each point represents a single taxon within that rank, numbers based on\nan average species copy number") +
	scale_x_discrete(breaks=c("kingdom", "phylum", "class", "order", "family", "genus", "species"),
									 labels=c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"))
```

* Bacteria have more copies than Archaea
* Even after correcting for number of genomes per species, there is wide variation in the number of rrn operons per taxonomic group