### 2020-10-16 exploratory/2020-10-15-asv-species-coverage.Rmd
Add table showing counts of most abundant species, closes #30

---
title: "Quantifying the coverage of ASVs within a species"
author: "Pat Schloss"
date: "10/15/2020"
output:
  github_document:
    html_preview: false
editor_options: 
  chunk_output_type: console
---

```{r, message=FALSE}
library(tidyverse)
library(knitr)
library(here)

metadata <- read_tsv(here("data/references/genome_id_taxonomy.tsv"),
										 col_types = cols(.default = col_character())) %>%
	mutate(strain = if_else(scientific_name == species,
													NA_character_,
													scientific_name)) %>%
	select(-scientific_name)

asv <- read_tsv(here("data/processed/rrnDB.count_tibble"),
								col_types = cols(.default = col_character(),
																 count = col_integer()))

metadata_asv <- inner_join(metadata, asv, by=c("genome_id" = "genome"))
```

### Does the number of ASVs per species increase with sampling effort? Does this vary by region within the 16S rRNA gene?

We've previously seen that the number of ASVs per copy of the 16S rRNA gene is about 0.6. This means that if a genome has 10 copies of the 16S rRNA gene, we would expect to see about 6 different versions of the gene within that genome. Across genomes from the same species, we know that the number of copies of the gene is pretty consistent. So, if we look at multiple genomes from the same species, will we see the same versions of the gene or will we see new versions of the gene? To answer this, we would like to look at the relationship between the number of ASVs found in a species per number of genomes in the species versus the number of genomes. Besides looking at the full length sequences, let's also look at the V4, V3-V4, and V4-V5 regions.

```{r}
#x = number of genomes for that species
#y = ratio of the number of ASVs per genome
# each point represents a different species
# each facet represents a different region

species_asvs <- metadata_asv %>% 
	select(genome_id, species, region, asv, count) %>%
	group_by(region, species) %>%
	summarize(n_genomes = n_distinct(genome_id),
						n_asvs = n_distinct(asv),
						n_rrns = sum(count) / n_genomes,
						asv_rate = n_asvs/n_genomes,
						.groups="drop")

region_labels <- c("V1-V9", "V4", "V3-V4", "V4-V5")
names(region_labels) <- c("v19", "v4", "v34", "v45")

species_asvs %>%
	ggplot(aes(x=n_genomes, y=asv_rate)) +
	geom_point(alpha=0.2) +
	geom_smooth() + 
	facet_wrap(facet="region",
						 nrow=2,
						 strip.position = "top",
						 scales="fixed",
						 labeller=labeller(region = region_labels)) +
	scale_x_log10()+
	scale_y_log10() +
	theme_classic()
```

```{r}
# species name, number of genomes, average number of rrns, number of ASVs across
# genomes for each region of 16S rRNA gene

count_table <- species_asvs %>%
	select(species, region, n_genomes, n_rrns, n_asvs) %>%	
	group_by(species) %>%
	mutate(n_genomes = max(n_genomes),
				 n_rrns = max(n_rrns)) %>%
	ungroup() %>% 
	pivot_wider(names_from=region, values_from=n_asvs) %>% 
	arrange(species)

#see also kableExtra
count_table %>%
	arrange(desc(n_genomes)) %>%
	top_n(n_genomes, n=10) %>%
	kable(caption = "Ten most commonly sequence species", digits=2)
```

### Conclusions

* V1-V9 continues to add significant number of ASVs as more genomes are sampled
from a species
* The sub regions seem to have plateaued indicating that perhaps we'll always
add more ASVs for a species
* Perhaps we really are splitting genomes and species too finely with ASVs
* Would like to look at individual genomes in more detail