### 2020-10-21 exploratory/2020-10-21-asv-taxa-overlap.Rmd
Create plot of overlap of ASVs across taxa with a rank, addresses #31

---
title: "Quantifying the overlap of ASVs between taxa"
author: "Pat Schloss"
date: "10/21/2020"
output:
  github_document:
    html_preview: false
editor_options: 
  chunk_output_type: inline
---

```{r, message=FALSE}
library(tidyverse)
library(here)
library(knitr)

metadata <- read_tsv(here("data/references/genome_id_taxonomy.tsv"),
										 col_types = cols(.default = col_character())) %>%
	mutate(strain = if_else(scientific_name == species,
													NA_character_,
													scientific_name)) %>%
	select(-scientific_name)

asv <- read_tsv(here("data/processed/rrnDB.count_tibble"),
								col_types = cols(.default = col_character(),
																 count = col_integer()))

metadata_asv <- inner_join(metadata, asv, by=c("genome_id" = "genome"))
```

### How often is the same ASV found in multiple taxa from the same rank?

Across the analyses, we've already seen that a single bacterial genome can have multiple ASVs and that depending on the species we are looking at, there may be as many ASVs as there are genome sequences for that species. To me this says that ASVs pose a risk at unnaturally splitting a species and even a genome into multiple taxonomic groupings! In my book, that's not good.

Today, I want to ask an equally important question: If I have an ASV, what's the probability that it is also found in another taxonomic group from the same rank? In other words, if I have an ASV from *Bacillus subtilis*, what's the probability of finding it in *Bacillus cereus*? Of course, it's probably more likely to find a *Bacillus subtilis* ASV in a more closely related organism like *Bacillus cereus* than *E. coli*. Perhaps we can control for relatedness in a future episode, but for today, I want to answer this question for any two taxa from the same rank.

```{r}
# metadata_asv - input data
overlap_data <- metadata_asv %>%
# - focus on taxonomic ranks - kingdom to species, asv, and region
	select(-genome_id, -count, -strain) %>%
# - make data frame tidy
	pivot_longer(cols=c(-region, -asv), names_to="rank", values_to="taxon") %>%
# - remove lines from the data where we don't have a taxonomy
	drop_na(taxon) %>%
# - remove redundant lines 
	distinct() %>%
# for each region and taxonomic rank, group by asvs
	group_by(region, rank, asv) %>%
# - for each asv - count the number of taxa
	summarize(n_taxa = n(), .groups="drop_last") %>%
# - count the number of asvs that appear in more than one taxa
	count(n_taxa) %>%
# - find the ratio of the number of asvs in more than one taxa to the total number of asvs
	summarize(overlap = sum((n_taxa > 1) * n) / sum(n), .groups="drop") %>%
	mutate(rank = factor(rank, levels=c("kingdom", "phylum", "class", "order", "family", "genus", "species"))) %>%
	mutate(overlap = 100*overlap)

# create a plot showing specificity at each taxonomic rank for each region
# x = taxonomic rank
# y = specificity or % of asvs found in more than one taxa
# geom = line plot
# different lines for each region of 16S rRNA gene
overlap_data %>%
	ggplot(aes(x=rank, y=overlap, group=region, color=region)) +
	geom_line()

overlap_data %>%
	filter(rank == "species") %>% kable
```

### Conclusions
* Sub regions are less specific at the species level than full-length. Still full-length has 3.6% overlap whereas the subregions are between 8.8 and 12.6% overlap
* This analysis does not control for uneven sampling of different species