### 2020-11-04 exploratory/2020-10-21-esv-taxa-overlap.Rmd

---
title: "Quantifying the overlap of ESVs between taxa"
author: "Pat Schloss"
date: "10/21/2020"
output:
  github_document:
    html_preview: false
editor_options:
  chunk_output_type: console
---

```{r, message=FALSE}
library(tidyverse)
library(here)
library(knitr)

metadata <- read_tsv(here("data/references/genome_id_taxonomy.tsv"),
										 col_types = cols(.default = col_character())) %>%
	mutate(strain = if_else(scientific_name == species,
													NA_character_,
													scientific_name)) %>%
	select(-scientific_name)

esv <- read_tsv(here("data/processed/rrnDB.esv.count_tibble"),
								col_types = cols(.default = col_character(),
																 count = col_integer()))

metadata_esv <- inner_join(metadata, esv, by=c("genome_id" = "genome"))
```

### How often is the same ESV found in multiple taxa from the same rank?

Across the analyses, we've already seen that a single bacterial genome can have multiple ESVs and that depending on the species we are looking at, there may be as many ESVs as there are genome sequences for that species. To me this says that ESVs pose a risk at unnaturally splitting a species and even a genome into multiple taxonomic groupings! In my book, that's not good.

Today, I want to ask an equally important question: If I have an ESV, what's the probability that it is also found in another taxonomic group from the same rank? In other words, if I have an ESV from *Bacillus subtilis*, what's the probability of finding it in *Bacillus cereus*? Of course, it's probably more likely to find a *Bacillus subtilis* ESV in a more closely related organism like *Bacillus cereus* than *E. coli*. Perhaps we can control for relatedness in a future episode, but for today, I want to answer this question for any two taxa from the same rank.

```{r}
# set RNG seed to Pat's birthday!
set.seed(19760620)

get_subsample_result <- function(threshold){
	# metadata_esv - species - genome_id
	subsample_species <- metadata_esv %>%
		select(genome_id, species) %>%
	# return the distinct/unique rows
		distinct() %>%
	# group_by the species
		group_by(species) %>%
	# slice_sample on each species for N genomes
		slice_sample(n=threshold) %>%
		ungroup()

	good_species <- subsample_species %>%
	# count number of genoems in each species
		count(species) %>%
	# return species that have N genomes / filter out species with fewer than N genomes
		filter(n == threshold) %>%
		select(species)

	# going back to original list of species/genomes and return genome_ids from
	# species with N genomes

	subsampled_genomes <- inner_join(subsample_species, good_species,  by="species") %>%
		select(genome_id)

	# metadata_esv - input data
	overlap_data <- inner_join(metadata_esv, subsampled_genomes, by="genome_id") %>%
	# - focus on taxonomic ranks - kingdom to species, esv, and region
		select(-genome_id, -count, -strain) %>%
	# - make data frame tidy
		pivot_longer(cols=c(-region, -esv), names_to="rank", values_to="taxon") %>%
	# - remove lines from the data where we don't have a taxonomy
		drop_na(taxon) %>%
	# - remove redundant lines
		distinct() %>%
	# for each region and taxonomic rank, group by esvs
		group_by(region, rank, esv) %>%
	# - for each esv - count the number of taxa
		summarize(n_taxa = n(), .groups="drop_last") %>%
	# - count the number of esvs that appear in more than one taxa
		count(n_taxa) %>%
	# - find the ratio of the number of esvs in more than one taxa to the total number of esvs
		summarize(overlap = sum((n_taxa > 1) * n) / sum(n), .groups="drop") %>%
		mutate(rank = factor(rank, levels=c("kingdom", "phylum", "class", "order", "family", "genus", "species"))) %>%
		mutate(overlap = 100*overlap)

	return(overlap_data)
}

subsample_iterations <- map_dfr(1:100, ~get_subsample_result(threshold=1), .id="iterations")

summary_overlap_data <- subsample_iterations %>%
	group_by(region, rank) %>%
	summarize(mean_overlap = mean(overlap),
						lci = quantile(overlap, prob=0.025),
						uci = quantile(overlap, prob=0.975),
						.groups = "drop")

# create a plot showing specificity at each taxonomic rank for each region
# x = taxonomic rank
# y = specificity or % of esvs found in more than one taxa
# geom = line plot
# different lines for each region of 16S rRNA gene

summary_overlap_data %>%
	ggplot(aes(x=rank, y=mean_overlap, group=region, color=region, ymin=lci, ymax=uci)) +
	geom_linerange(show.legend=FALSE) +
	geom_line()

summary_overlap_data %>%
	filter(rank == "species") %>% kable(digits=1)
```

### Conclusions
* Sub regions are less specific at every taxonomic rank than full-length sequences. Still full-length has 3.6% overlap whereas the sub-regions are between 10.2 and 15.0% overlap