### 2020-11-04 exploratory/2020-11-02-dominance-commonness-of-esvs.Rmd

---
title: "Quantifying the dominance and commoness of ESVs"
author: "Pat Schloss"
date: "11/02/2020"
output:
  github_document:
    html_preview: false
editor_options:
  chunk_output_type: console
---

```{r, message=FALSE}
library(tidyverse)
library(here)
library(knitr)

metadata <- read_tsv(here("data/references/genome_id_taxonomy.tsv"),
										 col_types = cols(.default = col_character())) %>%
	mutate(strain = if_else(scientific_name == species,
													NA_character_,
													scientific_name)) %>%
	select(-scientific_name)

esv <- read_tsv(here("data/processed/rrnDB.esv.count_tibble"),
								col_types = cols(.default = col_character(),
																 count = col_integer()))

metadata_esv <- inner_join(metadata, esv, by=c("genome_id" = "genome"))
```

maybe there are many ESVs but only a handful of them are dominant?

```{r}

# want the number of genomes and average number of 16S rRNA gene (rrn) copies per
# genome for each species

n_genomes_rrn_copies_per_species <-
	metadata_esv %>%
	# need the counts by region and by genome for each species
  select(region, genome_id, species, esv, count) %>%
	# group our ESVs by region and species
  group_by(region, species) %>%
  summarize(
  	# number of distinct genomes per region/species
  	n_genomes = n_distinct(genome_id),
  	# summing up the number of 16S rRNA copies across all genomes for
  	# a species and region and divide by number of genomes
  	mean_rrns = sum(count) / n_genomes,
  	.groups="drop"
  )

# set the minimum number of genomes per species
min_n_genomes_per_species <- 5

# What percentage of genomes have an ESV? This focuses on the most
# common ESV for each species. Dominance is the percentage of genomes that an
# ESV is found in. So if a ESV shows up in 20 of 25 genomes for a species, it
# dominance is 0.80.

esv_dominance_per_species <-

	metadata_esv %>%

	# want to do our analysis for the species level for each region of 16S will
	# want to group our genomes by ESVs and species to count the number of genomes
	# that each ESV appears in
  select(region, genome_id, species, esv, count) %>%
  group_by(region, species, esv) %>%

	# count the number of genomes that each ESV appears in for region and species
	summarize(n_genomes_found = n_distinct(genome_id), .groups="drop") %>%

	# bring in the number of genomes and rrn copies per species so we can scale
	# our number of genomes that each ESV appeared in to get percent dominance
  inner_join(., n_genomes_rrn_copies_per_species, by=c("region", "species")) %>%
  mutate(dominance = n_genomes_found / n_genomes) %>%

	# want the dominant ESV for each region and species
	# want the number of genomes for each species and region so that we can filter
	#   our data to focus on those species with more genomes than the value of
	#   threshold
  group_by(region, species) %>%
  summarize(
  	n_genomes=max(n_genomes),
  	max_dominance = max(dominance),
  	.groups="drop"
  	) %>%
  filter(n_genomes >= min_n_genomes_per_species)

esv_dominance_per_species %>%
  ggplot(aes(x=max_dominance, fill=region)) +
    geom_histogram(binwidth=0.05) +
    facet_wrap(facet="region") +
		labs(y="Number of species", x="Dominance of ESV per species")

# what percentage of species have a max dominance over 80%
esv_dominance_per_species %>%
	group_by(region) %>%
  summarize(fraction = sum(max_dominance > 0.80)/n())
```

```{r}
esv_commonness_per_species <- metadata_esv %>%

	# want to do our analysis for the species level for each region of 16S will
	# want to group our genomes by ESVs and species to count the number of genomes
	# that each ESV appears in
  select(region, genome_id, species, esv, count) %>%
  group_by(region, species, esv) %>%

	# instead of counting the number of genoems an ESV is found in, I want to
	# count the total number of times each ESV appears across the genomes within
	# a species
	summarize(n_esvs = sum(count), .groups="drop") %>%

	# bring in the number of genomes and rrn copies per species so we can scale
	# our number of times that each ESV appeared in each species to get percent
	# commonness
  inner_join(., n_genomes_rrn_copies_per_species, by=c("region", "species")) %>%
  mutate(commonness = n_esvs/n_genomes/mean_rrns) %>%

	# want the most common ESV for each region and species
	# want the number of genomes for each species and region so that we can filter
	#   our data to focus on those species with more genomes than the value of
	#   threshold
	group_by(region, species) %>%
  summarize(n_genomes=max(n_genomes),
    max_commonness = max(commonness),
    .groups="drop") %>%
  filter(n_genomes >= min_n_genomes_per_species)

esv_commonness_per_species %>%
  ggplot(aes(x=max_commonness, fill=region)) +
    geom_histogram(binwidth=0.05) +
    facet_wrap(facet="region") +
		labs(y="Number of species", x="Commonness of ESV per species")

# what percentage of species have an ESV that has a commonness over 80%
esv_commonness_per_species %>%
  group_by(region) %>%
  summarize(fraction = sum(max_commonness > 0.80)/n())
```

### Conclusions...
* Among the sub-regions, a majority of species have an ESV that is found in
  more than 80% of the genomes. For full length sequences, only 25% of ESVs
  are found across more than 80% of the genomes
* The most common ESVs (i.e. more than 80% of the rrn copies) account for
  46-74% of the genomes in the sub-regions and only 12% for the full length
  ESVs
* Underscores the problem that a single ESV is unlikely to be representative
  of the diversity of ESVs within the species