### 2020-11-24 exploratory/2020-11-24-threshold-to-drop-n-asvs.Rmd
Generate plots looking at E/ASVs per genome, closes #35

---
title: "Determine the threshold needed to get 1 ASV per genome"
author: "Pat Schloss"
date: "11/24/2020"
output:
  github_document:
    html_preview: false
editor_options:
  chunk_output_type: console
---

```{r, message=FALSE}
library(tidyverse)
library(here)
library(knitr)

set.seed(19760620)

metadata <- read_tsv(here("data/references/genome_id_taxonomy.tsv"),
										 col_types = cols(.default = col_character())) %>%
	select(genome_id, species) %>%
	group_by(species) %>% # Get one genome per species
	slice_sample(n=1) %>%
	ungroup()

easv <- read_tsv(here("data/processed/rrnDB.easv.count_tibble"),
								col_types = cols(.default = col_character(),
																 count = col_integer()))

metadata_easv <- inner_join(metadata, easv, by=c("genome_id" = "genome")) %>%
	mutate(threshold = recode(threshold, "esv" = "0.000"),
				 threshold = as.numeric(threshold))
```

### Overivew

We saw earlier that a problem with E/ASVs is the risk of splitting the operons from the same genome into multiple taxonomic groups. I would like to determine the threshold we should use so that 95% of the genomes are represented by a single E/ASV.
* Create a line plot of the 95th percentile for each threshold at different number of operons per genome. This should be faceted by region within the 16S rRNA gene.
* Create a line plot showing the 95th percentile line for number of E/ASVs per genome for each region when we only consider those genomes with 7 copies of the *rrn* operon

* Notes:
  * Should sample 1 genome per species
  * Is 95th percentile too strict?

```{r}
easvs_by_threshold_region <- metadata_easv %>%

# Aggregate data by species/genome, region, threshold, 
	group_by(region, threshold, genome_id) %>%

# Count the number of E/ASVs per grouping and the # rrns per genome
	summarize(n_rrns = sum(count), n_easvs = n(), .groups="drop") %>%

# Determine the 95th percentile for each region, threshold, and # rrns
	group_by(region, threshold, n_rrns) %>%
	summarize(upper_bound = quantile(n_easvs, prob=0.95), .groups="drop")

# Plot the...
#   * 95th percentile as a function of region,  threshold, #rrns
easvs_by_threshold_region %>%
	filter(threshold %in% c(0, 0.01, 0.02, 0.03)) %>%
	mutate(threshold = as.character(threshold)) %>%
	ggplot(aes(x=n_rrns, y=upper_bound, color=threshold)) +
	geom_line() +
	facet_wrap(~region, nrow=4)

#   * 95th percentile for the number of E/ASVs for those genomes with 7 copies 
#     of the rrn operon
easvs_by_threshold_region %>%
	filter(n_rrns == 4) %>%
	ggplot(aes(x=threshold, y=upper_bound, color=region)) +
	geom_line()
```

### Conclusions...
* Analysis depends on our comfort with uncertainty (prob=90%? 95?) and the number of rrn copies per genome
* Regardless of the the number of operons or region, we need a significantly higher threshold than ESV or even "traditional" ASVs are defined at. Frankly, 3% doesn't look so bad for this type of analysis.