### 2020-11-25 exploratory/2020-11-30-lumping-and-splitting.Rmd
Compare rate of lumping and splitting by threshold and region, closes #36

---
title: "What is the effect of distance thrshold on the lumping and splitting of bacterial species and genomes?"
author: "Pat Schloss"
date: "11/30/2020"
output:
  github_document:
    html_preview: false
editor_options:
  chunk_output_type: console
---

```{r, message=FALSE}
library(tidyverse)
library(here)
library(knitr)

set.seed(19760620)

metadata <- read_tsv(here("data/references/genome_id_taxonomy.tsv"),
										 col_types = cols(.default = col_character())) %>%
	select(genome_id, species) %>%
	group_by(species) %>% # Get one genome per species
	slice_sample(n=1) %>%
	ungroup()

easv <- read_tsv(here("data/processed/rrnDB.easv.count_tibble"),
								col_types = cols(.default = col_character(),
																 count = col_integer()))

metadata_easv <- inner_join(metadata, easv, by=c("genome_id" = "genome")) %>%
	mutate(threshold = recode(threshold, "esv" = "0.000"),
				 threshold = as.numeric(threshold))
```

### Overivew
Besides the risk of splitting a genome into multiple taxonomic groups, there's also the possibility that the same E/ASV can appear in multiple species. In otherwords, by using too broad of a threshold to define an ASV, we run the risk of lumping different species together (e.g. *Bacillus cereus*, *B. anthracis*, and *B. thuringiensis*). I would like to determine...

* How often is the same E/ASV found in the multiple species?
* What fraction of species have multiple E/ASVs?
* Create a plot showing the fraction of species with multiple E/ASVs and the fraction of E/ASVs that appear in multiple species as a function of the threshold used to define E/ASVs.

Notes:
* Determine for each region of the 16S rRNA gene
* Select one genome per species

```{r}
# Measuring degree of splitting...
# Determine fraction of genomes with more than one E/ASV by region and threshold
splitting_data <- metadata_easv %>%

# - group data by region, threshold, and genome_id
	group_by(region, threshold, genome_id) %>%
# - determine whether each genome has more than 1 E/ASV
	summarize(n_easvs = n_distinct(easv),
						is_split = n_easvs > 1, .groups="drop") %>%

# - for each region and threshold, determine the fraction of genomes with more
#   than one E/ASV
	group_by(region, threshold) %>%
	summarize(f_split = sum(is_split) / n(), .groups="drop")

# Measure degree of lumping...
# Determine the fraction of E/ASVs that are found in more than one
# genome/species
lumping_data <- metadata_easv %>%

# - group data by region, threshold, and E/ASV
	group_by(region, threshold, easv) %>%

# - determine whether each E/ASV is observed in more than 1 genome
	summarize(n_genomes = n_distinct(genome_id),
						is_lumped = n_genomes > 1, .groups="drop")  %>%

# - for each region and threshold, determine the fraction of E/ASVs that appear
#   in more than one genome
	group_by(region, threshold) %>%
	summarize(f_lumped = sum(is_lumped)/n(), .groups="drop")

# Join lumping and splitting data
lumping_splitting_data <- inner_join(splitting_data, lumping_data, by=c("region", "threshold")) %>%

# * Tidy so that we have columns for region, threshold, lumping/splitting, and
#   the degree of lumping and splitting
	pivot_longer(cols=c(f_split, f_lumped), names_to="method", values_to="fraction")

# * Plot degree of lumping and splitting as a function of threshold for each
#   region. Try faceting...
#   - By region with lumping/splitting as separate lines
lumping_splitting_data %>%
	ggplot(aes(x=threshold, y=fraction, color=method)) +
	geom_line() +
	facet_wrap(~region) # i like this

#   - By lumping/splitting with region as separate lines
lumping_splitting_data %>%
	ggplot(aes(x=threshold, y=fraction, color=region)) + 
	geom_line() +
	facet_wrap(~method)

#   - By lumping/splitting and by region
lumping_splitting_data %>%
	ggplot(aes(x=threshold, y=fraction)) +
	geom_line() +
	facet_grid(region~method)

lumping_splitting_data %>%
	ggplot(aes(x=threshold, y=fraction)) +
	geom_line() +
	facet_grid(method~region) # i like this
```

### Conclusions...
* As we increase the threshold, splitting drops and lumping increases
* Again, we need a decent sized threshold (>0.01) to reduce the level of splitting
* Would prefer to limit splitting over lumping because species designations are too squishy - human made rather bacterial made