### 2020-12-16 exploratory/2020-12-21-roc-curve.Rmd
Build ROC curves and update data, closes #37

---
title: "Balancing sensitivity and specificity by varying the ASV definition"
author: "Pat Schloss"
date: "12/21/2020"
output:
  github_document:
    html_preview: false
editor_options:
  chunk_output_type: console
---

```{r, message=FALSE}
library(tidyverse)
library(here)
library(knitr)

sensitivity_specificity <- read_tsv(here("data/processed/rrnDB.roc.tsv"),
													 col_types = cols(.default = col_double(),
													 								 region = col_character()
													 								 )
													 ) %>%
	mutate(sensitivity = true_pos / (true_pos + false_neg),
				 specificity = true_neg / (true_neg + false_pos)) %>%
	group_by(region, threshold) %>%
	summarize(sensitivity = median(sensitivity),
						specificity = median(specificity),
						.groups="drop") %>%
	select(region, threshold, sensitivity, specificity)
```

### Overivew

The sensitivity and specificity of assigning 16S rRNA gene sequences to ASVs relative to the genome/species have tradeoffs. We see these tradeoffs in whether ASVs from the same genome are split apart (i.e. sensitivity) or whether genomes from different species have the same ASV (i.e. specificity). As we've discussed, we'd like to prioritize minimizing the likelihood that we split genomes apart (i.e. high specificity), but minimizing lumping (i.e. low specificity). We can look at how both measures vary with the threshold used to define ASVs.

* sensitivity (true positive rate): true_pos / (true_pos + false_neg)
* specificity (true negative rate): true_neg / (true_neg + false_pos)

### Receiver operator characteristic curve (ROC curve)

```{r}
sensitivity_specificity %>%
	ggplot(aes(x=1-specificity, y=sensitivity, color=region)) + 
	geom_line() +
	theme_classic()
```

### What are the sensitivity and specificity at a 3% difference?

```{r}
three <- sensitivity_specificity %>%
	filter(threshold == 0.03)

three

sensitivity_specificity %>%
	ggplot(aes(x=1-specificity, y=sensitivity, color=region)) + 
	geom_abline(intercept=1, slope=-1, color="gray") +
	geom_line() +
	geom_point(data=three, aes(x=1-specificity, y=sensitivity, color=region)) +
	theme_classic()
```

### What threshold provides the equal balance of sensitivity and specificity?

```{r}
balance <- sensitivity_specificity %>%
	mutate(diff=abs(sensitivity - specificity)) %>%
	group_by(region) %>%
	summarize(min_diff = min(diff),
						threshold = threshold[which.min(diff)],
						sensitivity = sensitivity[which.min(diff)],
						specificity = specificity[which.min(diff)],
						.groups="drop"
						)

balance

sensitivity_specificity %>%
	ggplot(aes(x=1-specificity, y=sensitivity, color=region)) + 
	geom_abline(intercept=1, slope=-1, color="gray") +
	geom_line() +
	geom_point(data=balance, aes(x=1-specificity, y=sensitivity, color=region)) +
	theme_classic()

```

### What threshold provides a sensitivity and specificity closest to perfect classification?

```{r}
distance <- sensitivity_specificity %>%
	mutate(distance = sqrt((specificity - 1)^2 + (sensitivity - 1)^2)) %>%
	group_by(region) %>%
	summarize(min_distance = min(distance),
						threshold = threshold[which.min(distance)],
						sensitivity = sensitivity[which.min(distance)],
						specificity = specificity[which.min(distance)],
						.groups="drop"
						)

distance

sensitivity_specificity %>%
	ggplot(aes(x=1-specificity, y=sensitivity, color=region)) + 
	geom_abline(intercept=1, slope=-1, color="gray") +
	geom_line() +
	geom_point(data=distance, aes(x=1-specificity, y=sensitivity, color=region)) +
	theme_classic()
```

### Conclusions...
* We need more points going out to 0.10
* We need more points between percent thresholds (e.g. 0.025)
* Appears that "larger" thresholds provide the best balance and overall classification of ASVs relative to genomes/species