### 2021-02-05 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {}%
  {\par}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\hypertarget{amplicon-sequence-variants-should-not-replace-operational-taxonomic-units-in-marker-gene-data-analysis}{%
\section{Amplicon sequence variants should not replace operational
taxonomic units in marker-gene data
analysis}\label{amplicon-sequence-variants-should-not-replace-operational-taxonomic-units-in-marker-gene-data-analysis}}

\vspace{20mm}

\textbf{Running title:} ASVs vs.~OTUs

\vspace{20mm}

Patrick D. Schloss\({^\dagger}\)

\vspace{40mm}

\({\dagger}\) To whom corresponsdence should be addressed:

\href{mailto:pschloss@umich.edu}{pschloss@umich.edu}

Department of Microbiology \& Immunology

University of Michigan

Ann Arbor, MI 48109

\vspace{20mm}

\textbf{Observation Format}

\newpage

\hypertarget{abstract-250-words}{%
\subsection{Abstract (250 words)}\label{abstract-250-words}}

\hypertarget{importance-150-words}{%
\subsection{Importance (150 words)}\label{importance-150-words}}

\newpage

16S rRNA gene sequencing is a powerful technique for describing and
comparing microbial communities (1). Efforts to link 16S rRNA gene
sequences to taxonomic levels based on distance thresholds go back to at
least the 1990s. The distance-based thresholds that were developed and
are now widely used (3\%) were based on DNA-DNA hybridization approaches
that are not as precise as genome sequencing (2, 3). Instead, genome
sequencing technologies have suggested that the widely used 3\% distance
threshold to operationally define bacterial taxa is too coarse (4--6).
As an alternative to OTUs, amplicon sequencing variants (ASVs) have been
proposed as a way to adopt the thresholds suggested by genome sequencing
to microbial community analysis using 16S rRNA gene sequences (7--10).
ASVs are a unit of microbial community inference that do not cluster
sequences based on a distance-based threshold (11). However, most
bacterial genomes have more than 1 copy of the rrn operon and those
copies are not identical (12, 13). Therefore, using too fine a threshold
to identify OTUs creates the risk of splitting a single genome into
multiple bins and using too broad of a threshold to define OTUs creates
the risk of lumping together bacterial species into the same OTU. An
example of both is seen in the comparison of \emph{Staphylococcus
aureus} (NCTC 8325) and \emph{S. epidermidis} (ATCC 12228) where each
genome has 5 copies of the 16S rRNA gene. The 10 copies of the 16S rRNA
gene each have a different sequence and so if OTUs are defined based on
ASVs, each genome would be split into 5 OTUs. Conversely, if the copies
were clustered using a 3\% distance threshold all 10 copies would
cluster into the same OTU. The goal of this study was to quantify the
risk of splitting a single genome into multiple bins and the risk of
lumping together different bacterial species into the same bin.

To investigate the variation in the number of copies of the 16S rRNA
gene per genome as well as the intragenomic variation among copies of
the 16S rRNA gene, I obtained reference 16S rRNA sequences from the rrn
copy number database (rrnDB)(14). Among the \textbf{4,774} species
represented in the rrnDB there were \textbf{15,614} genomes. The median
number of rrn operson per species ranged between \textbf{1} (e.g.,
\textbf{\emph{Mycobacterium tuberculosis}}) and \textbf{19}
(\textbf{\emph{Metabacillus litoralis}}) copies of the rrn operon. As
the number of copies of the operon in a genome increased, the number of
variants of the 16S rRNA gene in each genome also increased. On average,
there were \textbf{1} variants per copy of the full length 16S rRNA gene
and an average of \textbf{0.26}, \textbf{0.33}, and \textbf{0.27}
variants when considering the V4, V3-V4, and V4-V5 regions of the gene,
respectively. Although a species tended to have a consistent number of
16S rRNA gene copies per genome, the number of total variants increased
with the number of genomes that were sampled (\textbf{Figure 1}). For
example, \textbf{\emph{Mycobacterium tuberculosis}} generally only had
\textbf{1} copy of the gene per genome, but across the \textbf{180}
genomes that have been sequenced there were \textbf{11} versions of the
gene. Similarly, a \emph{E. coli} genome typically had \textbf{7} copies
of the 16S rRNA gene with between \textbf{6} and \textbf{10} distinct
full length sequences per genome. Across the \textbf{958} \emph{E. coli}
genomes that have been sequenced, there were \textbf{1,013} different
variants of the gene. These observations highlight the risk of selecting
a threshold for defining units of inference that is too narrow because
it is possible to split a single genome into multiple units.

A method to avoid splitting a single genome into multiple units of
inference is to cluster together similar 16S rRNA gene sequences.
Therefore, I assessed the impact of the distance threshold used to
define clusters of 16S rRNA genes on the propensity to split a genome
into separate clusters. I observed that as the number of copies of the
\emph{rrn} operon increased, the distance threshold required to reduce
the ASVs in each genome to a single OTU increased (Figure 1). Among
species with 7 copies of the \emph{rrn} operon (e.g., \emph{E. coli}), I
found that a threshold of \textbf{5.5}\% was required to reduce full
length ASVs to a single OTU in 95\% of the species. Similarly,
thresholds of \textbf{2.5}, \textbf{4.0}, and \textbf{3.5}\% were
required for the V4, V3-V4, and V4-V5 regions, respectively. But, if a
3\% distance threshold was used, then ASVs from genomes containing fewer
than \textbf{5}, \textbf{8}, \textbf{6}, and \textbf{6} copies of the
\emph{rrn} operon would reliably be clustered into a single OTU for ASVs
from the V1-V9, V4, V3-V4, and V4-V5 regions, respectively.
Consequently, these results demonstrate that broad thresholds must be
used to avoid splitting different operons from the same genome into
separate clusters.

At broad thresholds multiple species could be represented by the same
OTU (\textbf{Figure 2}). Using ASVs, \textbf{3.6}\% of the species
shared a 16S rRNA gene sequence variant with another species when
considering full length sequences and \textbf{14.9}, \textbf{10.2}, and
\textbf{12.0}\% when considering the V4, V3-V4, and V4-V5 regions,
respectively. At the commonly used 3\% threshold, \textbf{25.2}\% of the
species shared an OTU when considering full length sequences and
\textbf{33.0}, \textbf{29.4}, and \textbf{32.2}\% when considering the
V4, V3-V4, and V4-V5 regions, respectively. Considering that species
designations are unevenly applied and reflect multiple biases, the risk
of splitting a genome into multiple OTUs more problematic than
clustering species together. Therefore, larger thresholds are advisable.

The results of this analysis demonstrate that there is a significant
risk of splitting single genomes into multiple bins if too fine of a
threshold is applied to defining an OTU. An ongoing problem for
amplicon-based studies is defining a meaningful taxonomic unit of
inference (11, 15, 16). Since there is no consensus definition for a
biological species concept (17, 18), microbiologists must accept that
how we have named bacterial species is biased and that taxonomic rules
are not applied in a consistent manner (e.g., (19)). This makes it more
challenging to attempt to fit a distance threshold to define an OTU
definition that matches a set of species names (20). Furthermore, the
16S rRNA gene does not evolve at the same rate across all bacterial
lineages (15), which limits the biological interpretation of a common
OTU definition. A distance-based definition of a taxonomic unit based on
16S rRNA gene or full genome sequences is, at best, operational and not
grounded in biological theory (15, 21--23). There is general agreement
in bacterial systematics that to classify something to a bacterial
species, you need phenotypic and genome sequence data (17--19). We are
asking too much of a short section of a bacterial genome to be able to
differentiate between species. It is difficult to defend a unit of
inference that would split a single genome into multiple taxonomic
units. It is not biologically plausible to entertain the possibility
that parts of a genome would have different ecologies. Although there
are multiple reasons that proponents of ASVs encourage their use, the
significant risk of splitting genomes is too high to warrant their use.

\textbf{Materials and Methods. (i) Data availability.} The 16S rRNA gene
sequences used in this study were obtained from the \emph{rrn}DB
(\url{https://rrndb.umms.med.umich.edu}; version 5.6, released November
8, 2019) (14). At the time of submission, this is the most current
version of the database. The \emph{rrn}DB obtained the curated 16S rRNA
gene sequences from the KEGG database, which ultimately obtained them
from NCBI's non-redundant RefSeq database. The \emph{rrn}DB provides
downloadable versions of the sequences with their taxonomy as determined
using the naive Bayesian classifier trained on the RDP reference
taxonomy. For some genomes this resulted in multiple classifications
since a genome's 16S rRNA gene sequences were not identical. Instead, I
mapped the RefSeq accession number for each genome in the database to
obtain a single taxonomy for each genome. Because strain names were not
consistently given to genomes across bacterial species, the strain level
designations were ignored.

\textbf{(ii) Definition of regions within 16S rRNA gene.} The full
length 16S rRNA gene sequences were aligned to a SILVA reference
alignment of the 16S rRNA gene (v138) using the mothur software package
(v. 1.XX) (24, 25). Regions of the 16S rRNA gene were selected because
of their use in the microbial ecology literature. Full length sequences
corresponded to \emph{E. coli} positions XX through XXXX, V4 to
positions XXX through XXX, V3-V4 to positions XXX through XXX, and V4-V5
to positions XXX through XXX.

\textbf{(iii) Controlling for uneven sampling of genomes by species.}
Because of the uneven distribution of genome sequences across species,
for the analysis of splitting genomes and lumping species I randomly
selected one genome for each species. The random selection was repeated
100 times. Analyses based on this randomization report the median of the
100 randomizations. The intraquartile range between randomizations was
typically less than XXXX. Because it was so small, confidence intervals
are not included in Figure 2.

\textbf{(iv) Reproducible data analysis.} The code to perform the
analysis in this manuscript and its hisotry are available as a git-based
version control repository on GitHub
(\url{https://github.com/pschloss/Schloss_rrnAnalysis_XXXX_2020}). The
analysis can be regenerated using a GNU Make-based workflow that made
use of built-in bash tools (v. 3.2.57), mothur (v. 1.XX), and R (v.
4.X.X). Within R, I used the tidyverse (v. 4.X.X), data.table (v.
4.X.X), Rcpp (v. 4.X.X), furrr (v. 4.X.X), and rmarkdown (v. 4.X.X)
packages. The conception and development of this analysis is available
as a playlist on the Riffomonas YouTube channel
(\url{https://www.youtube.com/playlist?list=PLmNrK_nkqBpKY3SZiivlIGvcLX-KHmfR8}).

\textbf{Acknowledgements.} I am grateful to Robert Hein and Thomas
Schmidt who maintain the rrnDB for their help in understanding the
curation of the database and for making the 16S rRNA gene sequences and
related metadata publicly available. I am also grateful to community
members who watched the serialized version of this analysis on YouTube
and provided their suggestions and questions.

This work was supported, in part, through grants from the NIH to PDS
(P30DK034933, U01AI124255, and R01CA215574).

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}

\noindent

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-Lane1985}{}%
1. Lane DJ, Pace B, Olsen GJ, Stahl DA, Sogin ML, Pace NR. 1985. Rapid
determination of 16S ribosomal RNA sequences for phylogenetic analyses.
Proceedings of the National Academy of Sciences 82:6955--6959
\url{https://doi.org/10.1073/pnas.82.20.6955}.

\leavevmode\hypertarget{ref-Stackebrandt1994}{}%
2. Stackebrandt E, Goebel BM. 1994. Taxonomic note: A place for DNA-DNA
reassociation and 16S rRNA sequence analysis in the present species
definition in bacteriology. International Journal of Systematic and
Evolutionary Microbiology 44:846--849
\url{https://doi.org/10.1099/00207713-44-4-846}.

\leavevmode\hypertarget{ref-Goris2007}{}%
3. Goris J, Konstantinidis KT, Klappenbach JA, Coenye T, Vandamme P,
Tiedje JM. 2007. DNA-DNA hybridization values and their relationship to
whole-genome sequence similarities. International Journal of Systematic
and Evolutionary Microbiology 57:81--91
\url{https://doi.org/10.1099/ijs.0.64483-0}.

\leavevmode\hypertarget{ref-RodriguezR2018}{}%
4. Rodriguez-R LM, Castro JC, Kyrpides NC, Cole JR, Tiedje JM,
Konstantinidis KT. 2018. How much do rRNA gene surveys underestimate
extant bacterial diversity? Applied and Environmental Microbiology
84:e00014--18 \url{https://doi.org/10.1128/aem.00014-18}.

\leavevmode\hypertarget{ref-Stackebrandt2006}{}%
5. Stackebrandt E, Ebers J. 2006. Taxonomic parameters revisited:
Tarnished gold standards. Microbiol Today 33:152--155.

\leavevmode\hypertarget{ref-Edgar2018}{}%
6. Edgar RC. 2018. Updating the 97\% identity threshold for 16S
ribosomal RNA OTUs. Bioinformatics 34:2371--2375
\url{https://doi.org/10.1093/bioinformatics/bty113}.

\leavevmode\hypertarget{ref-Edgar2016}{}%
7. Edgar RC. 2016. UNOISE2: Improved error-correction for illumina 16S
and its amplicon sequencing. bioRxiv
\url{https://doi.org/10.1101/081257}.

\leavevmode\hypertarget{ref-Amir2017}{}%
8. Amir A, McDonald D, Navas-Molina JA, Kopylova E, Morton JT, Zech Xu
Z, Kightley EP, Thompson LR, Hyde ER, Gonzalez A, Knight R. 2017. Deblur
rapidly resolves single-nucleotide community sequence patterns. mSystems
2:e00191--16 \url{https://doi.org/10.1128/mSystems.00191-16}.

\leavevmode\hypertarget{ref-Callahan2016}{}%
9. Callahan BJ, McMurdie PJ, Rosen MJ, Han AW, Johnson AJA, Holmes SP.
2016. DADA2: High-resolution sample inference from illumina amplicon
data. Nature Methods 13:581--583
\url{https://doi.org/10.1038/nmeth.3869}.

\leavevmode\hypertarget{ref-Eren2014}{}%
10. Eren AM, Morrison HG, Lescault PJ, Reveillaud J, Vineis JH, Sogin
ML. 2014. Minimum entropy decomposition: Unsupervised oligotyping for
sensitive partitioning of high-throughput marker gene sequences. The
ISME Journal 9:968--979 \url{https://doi.org/10.1038/ismej.2014.195}.

\leavevmode\hypertarget{ref-Callahan2017}{}%
11. Callahan BJ, McMurdie PJ, Holmes SP. 2017. Exact sequence variants
should replace operational taxonomic units in marker-gene data analysis.
The ISME Journal 11:2639--2643
\url{https://doi.org/10.1038/ismej.2017.119}.

\leavevmode\hypertarget{ref-Pei2010}{}%
12. Pei AY, Oberdorf WE, Nossa CW, Agarwal A, Chokshi P, Gerz EA, Jin Z,
Lee P, Yang L, Poles M, Brown SM, Sotero S, DeSantis T, Brodie E, Nelson
K, Pei Z. 2010. Diversity of 16S rRNA genes within individual
prokaryotic genomes. Applied and Environmental Microbiology
76:3886--3897 \url{https://doi.org/10.1128/aem.02953-09}.

\leavevmode\hypertarget{ref-Sun2013}{}%
13. Sun D-L, Jiang X, Wu QL, Zhou N-Y. 2013. Intragenomic heterogeneity
of 16S rRNA genes causes overestimation of prokaryotic diversity.
Applied and Environmental Microbiology 79:5962--5969
\url{https://doi.org/10.1128/aem.01282-13}.

\leavevmode\hypertarget{ref-Stoddard2014}{}%
14. Stoddard SF, Smith BJ, Hein R, Roller BRK, Schmidt TM. 2014. rrnDB:
Improved tools for interpreting rRNA gene abundance in bacteria and
archaea and a new foundation for future development. Nucleic Acids
Research 43:D593--D598 \url{https://doi.org/10.1093/nar/gku1201}.

\leavevmode\hypertarget{ref-Schloss2011}{}%
15. Schloss PD, Westcott SL. 2011. Assessing and improving methods used
in operational taxonomic unit-based approaches for 16S rRNA gene
sequence analysis. Applied and Environmental Microbiology 77:3219--3226
\url{https://doi.org/10.1128/aem.02810-10}.

\leavevmode\hypertarget{ref-Johnson2019}{}%
16. Johnson JS, Spakowicz DJ, Hong B-Y, Petersen LM, Demkowicz P, Chen
L, Leopold SR, Hanson BM, Agresta HO, Gerstein M, Sodergren E, Weinstock
GM. 2019. Evaluation of 16S rRNA gene sequencing for species and
strain-level microbiome analysis. Nature Communications 10:5029
\url{https://doi.org/10.1038/s41467-019-13036-1}.

\leavevmode\hypertarget{ref-Staley2006}{}%
17. Staley JT. 2006. The bacterial species dilemma and the
genomicphylogenetic species concept. Philosophical Transactions of the
Royal Society B: Biological Sciences 361:1899--1909
\url{https://doi.org/10.1098/rstb.2006.1914}.

\leavevmode\hypertarget{ref-Oren2013}{}%
18. Oren A, Garrity GM. 2013. Then and now: A systematic review of the
systematics of prokaryotes in the last 80~years. Antonie van Leeuwenhoek
106:43--56 \url{https://doi.org/10.1007/s10482-013-0084-1}.

\leavevmode\hypertarget{ref-Baltrus2016}{}%
19. Baltrus DA, McCann HC, Guttman DS. 2016. Evolution, genomics and
epidemiology ofPseudomonas syringae. Molecular Plant Pathology
18:152--168 \url{https://doi.org/10.1111/mpp.12506}.

\leavevmode\hypertarget{ref-Konstantinidis2005}{}%
20. Konstantinidis KT, Tiedje JM. 2005. Towards a genome-based taxonomy
for prokaryotes. Journal of Bacteriology 187:6258--6264
\url{https://doi.org/10.1128/jb.187.18.6258-6264.2005}.

\leavevmode\hypertarget{ref-Barco2020}{}%
21. Barco RA, Garrity GM, Scott JJ, Amend JP, Nealson KH, Emerson D.
2020. A genus definition for bacteria and archaea based on a standard
genome relatedness index. mBio 11:02475--19
\url{https://doi.org/10.1128/mbio.02475-19}.

\leavevmode\hypertarget{ref-Parks2020}{}%
22. Parks DH, Chuvochina M, Chaumeil P-A, Rinke C, Mussig AJ, Hugenholtz
P. 2020. A complete domain-to-species taxonomy for bacteria and archaea.
Nature Biotechnology 38:1079--1086
\url{https://doi.org/10.1038/s41587-020-0501-8}.

\leavevmode\hypertarget{ref-Yarza2014}{}%
23. Yarza P, Yilmaz P, Pruesse E, Glöckner FO, Ludwig W, Schleifer K-H,
Whitman WB, Euzéby J, Amann R, Rosselló-Móra R. 2014. Uniting the
classification of cultured and uncultured bacteria and archaea using 16S
rRNA gene sequences. Nature Reviews Microbiology 12:635--645
\url{https://doi.org/10.1038/nrmicro3330}.

\leavevmode\hypertarget{ref-Schloss2009}{}%
24. Schloss PD, Westcott SL, Ryabin T, Hall JR, Hartmann M, Hollister
EB, Lesniewski RA, Oakley BB, Parks DH, Robinson CJ, Sahl JW, Stres B,
Thallinger GG, Horn DJV, Weber CF. 2009. Introducing mothur:
Open-source, platform-independent, community-supported software for
describing and comparing microbial communities. Applied and
Environmental Microbiology 75:7537--7541
\url{https://doi.org/10.1128/aem.01541-09}.

\leavevmode\hypertarget{ref-Quast2012}{}%
25. Quast C, Pruesse E, Yilmaz P, Gerken J, Schweer T, Yarza P, Peplies
J, Glöckner FO. 2012. The SILVA ribosomal RNA gene database project:
Improved data processing and web-based tools. Nucleic Acids Research
41:D590--D596 \url{https://doi.org/10.1093/nar/gks1219}.
\end{cslreferences}

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

\newpage

\includegraphics{../figures/copy_number_threshold_plot.pdf}

\textbf{Figure 1. The distance threshold required to prevent the
splitting of genomes into multiple OTUs increases as the number of
\emph{rrn} operons in the genome increases.} Each line represents the
median distance threshold for each region of the 16S rRNA gene that is
required for 95\% of the species with the indicated numbrer of
\emph{rrn} operons to cluster their ASVs to a single OTU. The median
distance threshold was calculated across 100 randomizations in which one
genome was sampled from each species. Only those number of \emph{rrn}
operons that were found in more than 100 species are included.

\newpage

\includegraphics{../figures/lump_split.pdf}

\textbf{Figure 2. As the distance threshold used to define an OTU
increases, the fraction of genomes split into separate OTUs decreases
while the fraction of species that are merged into the same OTU
increases.} These data represent the median fractions for both
measurements across 100 randomizations. In each randomization, one
genome was sampled from each species.

\newpage

\includegraphics{../figures/esv_rate.pdf}

\textbf{Figure S1. The ratio of number of distinct ASVs per copy of the
\emph{rrn} operon increases for a species as the number of genomes
sampled increases.} Each point represents a different species and is
shaded to be 80\% transparent so that when points overlap they become
darker. The blue line represents a smoothed fit through the data.

\end{document}