### 2021-02-19 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {}%
  {\par}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\hypertarget{amplicon-sequence-variants-artificially-split-bacterial-genomes-into-separate-clusters}{%
\section{Amplicon sequence variants artificially split bacterial genomes
into separate
clusters}\label{amplicon-sequence-variants-artificially-split-bacterial-genomes-into-separate-clusters}}

\vspace{20mm}

\textbf{Running title:} ASVs artificially split bacterial genomes

\vspace{20mm}

Patrick D. Schloss\({^\dagger}\)

\vspace{40mm}

\({\dagger}\) To whom corresponsdence should be addressed:

\href{mailto:pschloss@umich.edu}{pschloss@umich.edu}

Department of Microbiology \& Immunology

University of Michigan

Ann Arbor, MI 48109

\vspace{20mm}

\textbf{Observation Format}

\newpage

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Amplicon sequencing variants (ASVs) have been proposed as an alternative
to operational taxonomic units (OTUs) for analyzing microbiomes. ASVs
have grown in popularity, in part, because of a desire to reflect a more
refined level of taxonomy because they do not cluster sequences based on
a distance-based threshold. However, ASVs and the use of overly narrow
thresholds to identify OTUs increases the risk of splitting a single
genome into separate clusters. I assessed this problem by analyzing the
intragenomic variation of 16S rRNA genes from the bacterial genomes
represented in a \emph{rrn} copy number database, which contained
\textbf{15,614} genomes from \textbf{4,774} species. As the number of
copies of the 16S rRNA gene increased in a genome, the number of ASVs
also increased. There was an average of \textbf{0.60} ASVs per copy of
the 16S rRNA gene, for full length 16S rRNA genes. It was necessary to
use a distance threshold of \textbf{5.5}\% to cluster the ASVs from the
same genome into a single OTU, for genomes with 7 copies of the 16S
rRNA, such as \emph{E. coli}. This research highlights the risk of
splitting a single bacterial genome into separate clusters when ASVs are
used to analyze 16S rRNA gene sequence data. Although there is also a
risk of clustering different species into the same OTU, the effects of
those risks on biological inferences are less than those from
artificially splitting a genome into separate ASVs and OTUs.

\hypertarget{importance}{%
\subsection{Importance}\label{importance}}

16S rRNA gene sequencing has propelled significant interest into
host-associated and environmental microbiomes. There is a tension
between trying to classify 16S rRNA gene sequences to increasingly lower
taxonomic levels and the reality that those levels were defined using
more sequence and physiological information than is available from a
fragment of the 16S rRNA gene. Furthermore, naming of bacterial taxa
reflects the biases of those who name them. One motivation for the
recent push to adopt ASVs in place of OTUs in microbiome analyses is to
allow researchers to perform their analyes at the finest possible level
that reflects species-level taxonomy. The current research is
significant because it quantifies the risk of artificially splitting
bacterial genomes into separate clusters. Far from providing a better
represenation of microbiomes, the ASV approach could lead to conflicting
inferences about the ecology of different clusters from the same genome.

\newpage

16S rRNA gene sequencing is a powerful technique for describing and
comparing microbial communities (1). Efforts to link 16S rRNA gene
sequences to taxonomic levels based on distance thresholds go back to at
least the 1990s. The distance-based thresholds that were developed and
are now widely used (i.e., 3\%) were based on DNA-DNA hybridization
approaches that are not as precise as genome sequencing (2, 3). Instead,
genome sequencing technologies have suggested that the widely used 3\%
distance threshold to operationally define bacterial taxa is too coarse
(4--6). As an alternative to OTUs, amplicon sequencing variants (ASVs)
have been proposed as a way to adopt the thresholds suggested by genome
sequencing to microbial community analysis using 16S rRNA gene sequences
(7--10). ASVs are a unit of microbial community inference that do not
cluster sequences based on a distance-based threshold (11). However,
most bacterial genomes have more than 1 copy of the \emph{rrn} operon
and those copies are not identical (12, 13). Therefore, using too fine a
threshold to identify OTUs creates the risk of splitting a single genome
into multiple bins and using too broad of a threshold to define OTUs
creates the risk of lumping together bacterial species into the same
OTU. An example of both is seen in the comparison of
\emph{Staphylococcus aureus} (NCTC 8325) and \emph{S. epidermidis} (ATCC
12228) where each genome has 5 copies of the 16S rRNA gene. Each of the
10 copies of the 16S rRNA gene in these two genomes were distinct and
would yield 10 ASVs. Conversely, if the copies were clustered using a
3\% distance threshold, then all 10 ASVs would cluster into the same
OTU. The goal of this study was to quantify the risk of splitting a
single genome into multiple clusters and the risk of lumping together
different bacterial species into the same cluster.

To investigate the variation in the number of copies of the 16S rRNA
gene per genome as well as the intragenomic variation among copies of
the 16S rRNA gene, I obtained reference 16S rRNA sequences from the
\emph{rrn} copy number database (\emph{rrn}DB)(14). Among the
\textbf{4,774} species represented in the \emph{rrn}DB there were
\textbf{15,614} genomes. The median number of \emph{rrn} operons per
species ranged between \textbf{1} (e.g., \textbf{\emph{Mycobacterium
tuberculosis}}) and \textbf{19} (\textbf{\emph{Metabacillus litoralis}})
copies of the \emph{rrn} operon. As the number of copies of the operon
in a genome increased, the number of variants of the 16S rRNA gene in
each genome also increased. On average, there were \textbf{0.60}
variants per copy of the full length 16S rRNA gene and an average of
\textbf{0.26}, \textbf{0.33}, and \textbf{0.27} variants when
considering the V4, V3-V4, and V4-V5 regions of the gene, respectively.
Although a species tended to have a consistent number of 16S rRNA gene
copies per genome, the number of total variants increased with the
number of genomes that were sampled (\textbf{Figure 1}). For example,
\textbf{\emph{Mycobacterium tuberculosis}} generally only had \textbf{1}
copy of the gene per genome, but across the \textbf{180} genomes in the
\emph{rrn}DB there were \textbf{11} versions of the gene. Similarly, a
\emph{E. coli} genome typically had \textbf{7} copies of the 16S rRNA
gene with between \textbf{6} and \textbf{10} distinct full length
sequences per genome. Across the \textbf{958} \emph{E. coli} genomes in
the \emph{rrn}DB, there were \textbf{1,013} different variants of the
gene. These observations highlight the risk of selecting a threshold for
defining clusters that is too narrow because it is possible to split a
single genome into multiple clusters.

A method to avoid splitting a single genome into multiple clusters is to
cluster 16S rRNA gene sequences together based on their similarity to
each other. Therefore, I assessed the impact of the distance threshold
used to define clusters of 16S rRNA genes on the propensity to split a
genome into separate clusters. I observed that as the number of copies
of the \emph{rrn} operon increased, the distance threshold required to
reduce the ASVs in each genome to a single OTU increased (Figure 1).
Among species with 7 copies of the \emph{rrn} operon (e.g., \emph{E.
coli}), I found that a threshold of \textbf{5.5}\% was required to
reduce full length ASVs to a single OTU in 95\% of the species.
Similarly, thresholds of \textbf{2.5}, \textbf{4.0}, and \textbf{3.5}\%
were required for the V4, V3-V4, and V4-V5 regions, respectively. But,
if a 3\% distance threshold was used, then ASVs from genomes containing
fewer than \textbf{5}, \textbf{8}, \textbf{6}, and \textbf{6} copies of
the \emph{rrn} operon would reliably be clustered into a single OTU for
ASVs from the V1-V9, V4, V3-V4, and V4-V5 regions, respectively.
Consequently, these results demonstrate that broad thresholds must be
used to avoid splitting different operons from the same genome into
separate clusters. At broad thresholds multiple species could be
represented by the same OTU (\textbf{Figure 2}). Using ASVs,
\textbf{3.6}\% of the species shared a 16S rRNA gene sequence variant
with another species when considering full length sequences and
\textbf{14.9}, \textbf{10.2}, and \textbf{12.0}\% when considering the
V4, V3-V4, and V4-V5 regions, respectively. At the commonly used 3\%
threshold, \textbf{25.2}\% of the species shared an OTU with another
speecies when considering full length sequences and \textbf{33.0},
\textbf{29.4}, and \textbf{32.2}\% when considering the V4, V3-V4, and
V4-V5 regions, respectively. Considering that species designations are
unevenly applied and reflect multiple human-imposed biases, the risk of
splitting a genome into multiple OTUs more problematic than clustering
species together. Therefore, larger thresholds are advisable.

The results of this analysis demonstrate that there is a significant
risk of splitting a single genome into multiple clusters if too fine of
a threshold is applied to defining an OTU. An ongoing problem for
amplicon-based studies is defining a meaningful taxonomic unit (11, 15,
16). Since there is no consensus definition for a biological species
concept (17, 18), microbiologists must accept that how we have named
bacterial species is biased and that taxonomic rules are not applied in
a consistent manner (e.g., (19)). This makes it impossible to fit a
distance threshold to define an OTU definition that matches a set of
species names (20). Furthermore, the 16S rRNA gene does not evolve at
the same rate across all bacterial lineages (15), which limits the
biological interpretation of a common OTU definition. A distance-based
definition of a taxonomic unit based on 16S rRNA gene or full genome
sequences is, at best, operational and not grounded in biological theory
(15, 21--23). There is general agreement in bacterial systematics that
to classify something to a bacterial species, phenotypic and genome
sequence data are needed (17--19). A short section of a bacterial genome
simply cannot differentiate between species. Moreover, it is difficult
to defend a clustering approach that would split a single genome into
multiple taxonomic units. It is not biologically plausible to entertain
the possibility that parts of a genome would have different ecologies.
Although there are multiple reasons that proponents favor ASVs, the
significant risk of artificially splitting genomes into separate
clusters is too high to warrant their use.

\textbf{Materials and Methods. (i) Data availability.} The 16S rRNA gene
sequences used in this study were obtained from the \emph{rrn}DB
(\url{https://rrndb.umms.med.umich.edu}; version 5.6, released November
8, 2019) (14). At the time of submission, this was the most current
version of the database. The \emph{rrn}DB obtained the curated 16S rRNA
gene sequences from the KEGG database, which ultimately obtained them
from NCBI's non-redundant RefSeq database. The \emph{rrn}DB provided
downloadable versions of the sequences with their taxonomy as determined
using the naive Bayesian classifier trained on the RDP reference
taxonomy. For some genomes this resulted in multiple classifications
since a genome's 16S rRNA gene sequences were not identical. Instead, I
mapped the RefSeq accession number for each genome in the database to
obtain a single taxonomy for each genome. Because strain names were not
consistently given to genomes across bacterial species, the strain level
designations were ignored.

\textbf{(ii) Definition of regions within 16S rRNA gene.} The full
length 16S rRNA gene sequences were aligned to a SILVA reference
alignment of the 16S rRNA gene (v. 138) using the mothur software
package (v. 1.44.2) (24, 25). Regions of the 16S rRNA gene were selected
because of their use in the microbial ecology literature. Full length
sequences corresponded to \emph{E. coli} str. K-12 substr. MG1655
(NC\_000913) positions 28 through 1491, V4 to positions 534 through 786,
V3-V4 to positions 358 through 786, and V4-V5 to positions 534 through
908.

\textbf{(iii) Controlling for uneven sampling of genomes by species.}
Because of the uneven distribution of genome sequences across species,
for the analysis of splitting genomes and lumping species I randomly
selected one genome for each species. The random selection was repeated
100 times. Analyses based on this randomization reported the median of
the 100 randomizations. The intraquartile range between randomizations
was less than XXXX. Because the range was so small, the confidence
intervals were smaller than the thickness of the lines in Figures 1 and
2 and were not included.

\textbf{(iv) Reproducible data analysis.} The code to perform the
analysis in this manuscript and its history are available as a git-based
version control repository on GitHub
(\url{https://github.com/SchlossLab/Schloss_rrnAnalysis_mSphere_2021}).
The analysis can be regenerated using a GNU Make-based workflow that
made use of built-in bash tools (v. 3.2.57), mothur (v. 1.44.2), and R
(v. \textbf{4.0.3}). Within R, I used the tidyverse (v. \textbf{1.3.0}),
data.table (v. \textbf{1.13.2}), Rcpp (v. packageVersion(``Rcpp'')),
furrr (v. packageVersion(``furrr'')), and rmarkdown (v.
packageVersion(``rmarkdown'')) packages. The conception and development
of this analysis is available as a playlist on the Riffomonas YouTube
channel
(\url{https://youtube.com/playlist?list=PLmNrK_nkqBpL7m_tyWdQgdyurerttCsPY}).

\textbf{Acknowledgements.} I am grateful to Robert Hein and Thomas
Schmidt, who maintain the \emph{rrn}DB, for their help in understanding
the curation of the database and for making the 16S rRNA gene sequences
and related metadata publicly available. I am also grateful to community
members who watched the serialized version of this analysis on YouTube
and provided suggestions and questions over the course of the
development of this project.

This work was supported, in part, through grants from the NIH to PDS
(P30DK034933, U01AI124255, and R01CA215574).

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}

\noindent

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-Lane1985}{}%
1. Lane DJ, Pace B, Olsen GJ, Stahl DA, Sogin ML, Pace NR. 1985. Rapid
determination of 16S ribosomal RNA sequences for phylogenetic analyses.
Proceedings of the National Academy of Sciences 82:6955--6959
\url{https://doi.org/10.1073/pnas.82.20.6955}.

\leavevmode\hypertarget{ref-Stackebrandt1994}{}%
2. Stackebrandt E, Goebel BM. 1994. Taxonomic note: A place for DNA-DNA
reassociation and 16S rRNA sequence analysis in the present species
definition in bacteriology. International Journal of Systematic and
Evolutionary Microbiology 44:846--849
\url{https://doi.org/10.1099/00207713-44-4-846}.

\leavevmode\hypertarget{ref-Goris2007}{}%
3. Goris J, Konstantinidis KT, Klappenbach JA, Coenye T, Vandamme P,
Tiedje JM. 2007. DNA-DNA hybridization values and their relationship to
whole-genome sequence similarities. International Journal of Systematic
and Evolutionary Microbiology 57:81--91
\url{https://doi.org/10.1099/ijs.0.64483-0}.

\leavevmode\hypertarget{ref-RodriguezR2018}{}%
4. Rodriguez-R LM, Castro JC, Kyrpides NC, Cole JR, Tiedje JM,
Konstantinidis KT. 2018. How much do rRNA gene surveys underestimate
extant bacterial diversity? Applied and Environmental Microbiology
84:e00014--18 \url{https://doi.org/10.1128/aem.00014-18}.

\leavevmode\hypertarget{ref-Stackebrandt2006}{}%
5. Stackebrandt E, Ebers J. 2006. Taxonomic parameters revisited:
Tarnished gold standards. Microbiol Today 33:152--155.

\leavevmode\hypertarget{ref-Edgar2018}{}%
6. Edgar RC. 2018. Updating the 97\% identity threshold for 16S
ribosomal RNA OTUs. Bioinformatics 34:2371--2375
\url{https://doi.org/10.1093/bioinformatics/bty113}.

\leavevmode\hypertarget{ref-Edgar2016}{}%
7. Edgar RC. 2016. UNOISE2: Improved error-correction for illumina 16S
and its amplicon sequencing. bioRxiv
\url{https://doi.org/10.1101/081257}.

\leavevmode\hypertarget{ref-Amir2017}{}%
8. Amir A, McDonald D, Navas-Molina JA, Kopylova E, Morton JT, Zech Xu
Z, Kightley EP, Thompson LR, Hyde ER, Gonzalez A, Knight R. 2017. Deblur
rapidly resolves single-nucleotide community sequence patterns. mSystems
2:e00191--16 \url{https://doi.org/10.1128/mSystems.00191-16}.

\leavevmode\hypertarget{ref-Callahan2016}{}%
9. Callahan BJ, McMurdie PJ, Rosen MJ, Han AW, Johnson AJA, Holmes SP.
2016. DADA2: High-resolution sample inference from illumina amplicon
data. Nature Methods 13:581--583
\url{https://doi.org/10.1038/nmeth.3869}.

\leavevmode\hypertarget{ref-Eren2014}{}%
10. Eren AM, Morrison HG, Lescault PJ, Reveillaud J, Vineis JH, Sogin
ML. 2014. Minimum entropy decomposition: Unsupervised oligotyping for
sensitive partitioning of high-throughput marker gene sequences. The
ISME Journal 9:968--979 \url{https://doi.org/10.1038/ismej.2014.195}.

\leavevmode\hypertarget{ref-Callahan2017}{}%
11. Callahan BJ, McMurdie PJ, Holmes SP. 2017. Exact sequence variants
should replace operational taxonomic units in marker-gene data analysis.
The ISME Journal 11:2639--2643
\url{https://doi.org/10.1038/ismej.2017.119}.

\leavevmode\hypertarget{ref-Pei2010}{}%
12. Pei AY, Oberdorf WE, Nossa CW, Agarwal A, Chokshi P, Gerz EA, Jin Z,
Lee P, Yang L, Poles M, Brown SM, Sotero S, DeSantis T, Brodie E, Nelson
K, Pei Z. 2010. Diversity of 16S rRNA genes within individual
prokaryotic genomes. Applied and Environmental Microbiology
76:3886--3897 \url{https://doi.org/10.1128/aem.02953-09}.

\leavevmode\hypertarget{ref-Sun2013}{}%
13. Sun D-L, Jiang X, Wu QL, Zhou N-Y. 2013. Intragenomic heterogeneity
of 16S rRNA genes causes overestimation of prokaryotic diversity.
Applied and Environmental Microbiology 79:5962--5969
\url{https://doi.org/10.1128/aem.01282-13}.

\leavevmode\hypertarget{ref-Stoddard2014}{}%
14. Stoddard SF, Smith BJ, Hein R, Roller BRK, Schmidt TM. 2014. rrnDB:
Improved tools for interpreting rRNA gene abundance in bacteria and
archaea and a new foundation for future development. Nucleic Acids
Research 43:D593--D598 \url{https://doi.org/10.1093/nar/gku1201}.

\leavevmode\hypertarget{ref-Schloss2011}{}%
15. Schloss PD, Westcott SL. 2011. Assessing and improving methods used
in operational taxonomic unit-based approaches for 16S rRNA gene
sequence analysis. Applied and Environmental Microbiology 77:3219--3226
\url{https://doi.org/10.1128/aem.02810-10}.

\leavevmode\hypertarget{ref-Johnson2019}{}%
16. Johnson JS, Spakowicz DJ, Hong B-Y, Petersen LM, Demkowicz P, Chen
L, Leopold SR, Hanson BM, Agresta HO, Gerstein M, Sodergren E, Weinstock
GM. 2019. Evaluation of 16S rRNA gene sequencing for species and
strain-level microbiome analysis. Nature Communications 10:5029
\url{https://doi.org/10.1038/s41467-019-13036-1}.

\leavevmode\hypertarget{ref-Staley2006}{}%
17. Staley JT. 2006. The bacterial species dilemma and the
genomicphylogenetic species concept. Philosophical Transactions of the
Royal Society B: Biological Sciences 361:1899--1909
\url{https://doi.org/10.1098/rstb.2006.1914}.

\leavevmode\hypertarget{ref-Oren2013}{}%
18. Oren A, Garrity GM. 2013. Then and now: A systematic review of the
systematics of prokaryotes in the last 80~years. Antonie van Leeuwenhoek
106:43--56 \url{https://doi.org/10.1007/s10482-013-0084-1}.

\leavevmode\hypertarget{ref-Baltrus2016}{}%
19. Baltrus DA, McCann HC, Guttman DS. 2016. Evolution, genomics and
epidemiology ofPseudomonas syringae. Molecular Plant Pathology
18:152--168 \url{https://doi.org/10.1111/mpp.12506}.

\leavevmode\hypertarget{ref-Konstantinidis2005}{}%
20. Konstantinidis KT, Tiedje JM. 2005. Towards a genome-based taxonomy
for prokaryotes. Journal of Bacteriology 187:6258--6264
\url{https://doi.org/10.1128/jb.187.18.6258-6264.2005}.

\leavevmode\hypertarget{ref-Barco2020}{}%
21. Barco RA, Garrity GM, Scott JJ, Amend JP, Nealson KH, Emerson D.
2020. A genus definition for bacteria and archaea based on a standard
genome relatedness index. mBio 11:02475--19
\url{https://doi.org/10.1128/mbio.02475-19}.

\leavevmode\hypertarget{ref-Parks2020}{}%
22. Parks DH, Chuvochina M, Chaumeil P-A, Rinke C, Mussig AJ, Hugenholtz
P. 2020. A complete domain-to-species taxonomy for bacteria and archaea.
Nature Biotechnology 38:1079--1086
\url{https://doi.org/10.1038/s41587-020-0501-8}.

\leavevmode\hypertarget{ref-Yarza2014}{}%
23. Yarza P, Yilmaz P, Pruesse E, Glöckner FO, Ludwig W, Schleifer K-H,
Whitman WB, Euzéby J, Amann R, Rosselló-Móra R. 2014. Uniting the
classification of cultured and uncultured bacteria and archaea using 16S
rRNA gene sequences. Nature Reviews Microbiology 12:635--645
\url{https://doi.org/10.1038/nrmicro3330}.

\leavevmode\hypertarget{ref-Schloss2009}{}%
24. Schloss PD, Westcott SL, Ryabin T, Hall JR, Hartmann M, Hollister
EB, Lesniewski RA, Oakley BB, Parks DH, Robinson CJ, Sahl JW, Stres B,
Thallinger GG, Horn DJV, Weber CF. 2009. Introducing mothur:
Open-source, platform-independent, community-supported software for
describing and comparing microbial communities. Applied and
Environmental Microbiology 75:7537--7541
\url{https://doi.org/10.1128/aem.01541-09}.

\leavevmode\hypertarget{ref-Quast2012}{}%
25. Quast C, Pruesse E, Yilmaz P, Gerken J, Schweer T, Yarza P, Peplies
J, Glöckner FO. 2012. The SILVA ribosomal RNA gene database project:
Improved data processing and web-based tools. Nucleic Acids Research
41:D590--D596 \url{https://doi.org/10.1093/nar/gks1219}.
\end{cslreferences}

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

\newpage

\includegraphics{../figures/copy_number_threshold_plot.pdf}

\textbf{Figure 1. The distance threshold required to prevent the
splitting of genomes into multiple OTUs increased as the number of
\emph{rrn} operons in the genome increased.} Each line represents the
median distance threshold for each region of the 16S rRNA gene that is
required for 95\% of the species with the indicated numbrer of
\emph{rrn} operons to cluster their ASVs to a single OTU. The median
distance threshold was calculated across 100 randomizations in which one
genome was sampled from each species. Only those number of \emph{rrn}
operons that were found in more than 100 species are included.

\newpage

\includegraphics{../figures/lump_split.pdf}

\textbf{Figure 2. As the distance threshold used to define an OTU
increased, the fraction of genomes split into separate OTUs decreased
while the fraction of species that were merged into the same OTU
increases.} These data represent the median fractions for both
measurements across 100 randomizations. In each randomization, one
genome was sampled from each species.

\newpage

\includegraphics{../figures/esv_rate.pdf}

\textbf{Figure S1. The ratio of number of distinct ASVs per copy of the
\emph{rrn} operon increased for a species as the number of genomes
sampled increased.} Each point represents a different species and was
shaded to be 80\% transparent so that when points overlap they become
darker. The blue line represents a smoothed fit through the data.

\end{document}