### 2021-06-23 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\raggedright
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {}%
  {\par}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\hypertarget{amplicon-sequence-variants-artificially-split-bacterial-genomes-into-separate-clusters}{%
\section{Amplicon sequence variants artificially split bacterial genomes
into separate
clusters}\label{amplicon-sequence-variants-artificially-split-bacterial-genomes-into-separate-clusters}}

\vspace{20mm}

\textbf{Running title:} ASVs artificially split bacterial genomes

\vspace{20mm}

Patrick D. Schloss\({^\dagger}\)

\vspace{40mm}

\({\dagger}\) To whom corresponsdence should be addressed:

\href{mailto:pschloss@umich.edu}{pschloss@umich.edu}

Department of Microbiology \& Immunology

University of Michigan

Ann Arbor, MI 48109

\vspace{20mm}

\textbf{Observation Format}

\newpage

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Amplicon sequencing variants (ASVs) have been proposed as an alternative
to operational taxonomic units (OTUs) for analyzing microbial
communities. ASVs have grown in popularity, in part, because of a desire
to reflect a more refined level of taxonomy since they do not cluster
sequences based on a distance-based threshold. However, ASVs and the use
of overly narrow thresholds to identify OTUs increase the risk of
splitting a single genome into separate clusters. To assess this risk, I
analyzed the intragenomic variation of 16S rRNA genes from the bacterial
genomes represented in a \emph{rrn} copy number database, which
contained 20,427 genomes from 5,972 species. As the number of copies of
the 16S rRNA gene increased in a genome, the number of ASVs also
increased. There was an average of 0.58 ASVs per copy of the 16S rRNA
gene for full length 16S rRNA genes. It was necessary to use a distance
threshold of 5.25\% to cluster full length ASVs from the same genome
into a single OTU with 95\% confidence for genomes with 7 copies of the
16S rRNA, such as \emph{E. coli}. This research highlights the risk of
splitting a single bacterial genome into separate clusters when ASVs are
used to analyze 16S rRNA gene sequence data. Although there is also a
risk of clustering ASVs from different species into the same OTU when
using broad distance thresholds, those risks are of less concern than
artificially splitting a genome into separate ASVs and OTUs.

\hypertarget{importance}{%
\subsection{Importance}\label{importance}}

16S rRNA gene sequencing has engendered significant interest in studying
microbial communities. There has been a tension between trying to
classify 16S rRNA gene sequences to increasingly lower taxonomic levels
and the reality that those levels were defined using more sequence and
physiological information than is available from a fragment of the 16S
rRNA gene. Furthermore, naming of bacterial taxa reflects the biases of
those who name them. One motivation for the recent push to adopt ASVs in
place of OTUs in microbial community analyses is to allow researchers to
perform their analyes at the finest possible level that reflects
species-level taxonomy. The current research is significant because it
quantifies the risk of artificially splitting bacterial genomes into
separate clusters. Far from providing a better represenation of
bacterial taxonomy and biology, the ASV approach can lead to conflicting
inferences about the ecology of different ASVs from the same genome.

\newpage

16S rRNA gene sequencing is a powerful technique for describing and
comparing microbial communities (1). Efforts to link 16S rRNA gene
sequences to taxonomic levels based on distance thresholds date to at
least the 1990s. The distance-based threshold that was developed and is
now widely used was based on DNA-DNA hybridization approaches that are
not as precise as genome sequencing (2, 3). Instead, genome sequencing
technologies have suggested that the widely used 3\% distance threshold
to operationally define bacterial taxa is too coarse (4--6). As an
alternative to operational taxonomic units (OTUs), amplicon sequencing
variants (ASVs) have been proposed as a way to adopt the thresholds
suggested by genome sequencing to microbial community analysis using 16S
rRNA gene sequences (7--10). It is widely understood that individual
bacterial genomes often have multiple 16S rRNA genes that are not
identical and that a 16S rRNA gene sequence could be found with
different versions of the sequence in different genomes (11, 12). This
could lead to the problem that ASVs and using too fine a threshold to
identify OTUs could split a single genome into multiple clusters.
Proponents of ASVs minimize concerns that most bacterial genomes have
more than one copy of the \emph{rrn} operon and that those copies are
not identical (6, 13). Conversely, using too broad of a threshold to
define OTUs could cluster together multiple bacterial species into the
same OTU. An example of both is seen in the comparison of
\emph{Staphylococcus aureus} (NCTC 8325) and \emph{S. epidermidis} (ATCC
12228) where each genome has 5 copies of the 16S rRNA gene. Each of the
10 copies of the 16S rRNA gene in these two genomes is distinct and
represent 10 ASVs. Conversely, if the copies were clustered using a 3\%
distance threshold, then all 10 ASVs would cluster into the same OTU.
The goal of this study was to quantify the tradeoff of splitting a
single genome into multiple clusters and the risk of clustering
different bacterial species into the same cluster when using ASVs and
various OTU definitions.

To investigate the variation in the number of copies of the 16S rRNA
gene per genome and the intragenomic variation among copies of the 16S
rRNA gene, I obtained 16S rRNA sequences from the \emph{rrn} copy number
database (\emph{rrn}DB)(14). Among the 5,972 species represented in the
\emph{rrn}DB there were 20,427 genomes. The median \emph{rrn} copy
number per species ranged between 1 (e.g., \emph{Mycobacterium
tuberculosis}) and 19 (\emph{Metabacillus litoralis}). As the \emph{rrn}
copy number for a genome increased, the number of variants of the 16S
rRNA gene in each genome also increased. On average, there were 0.58
variants per copy of the full length 16S rRNA gene and an average of
0.32, 0.25, and 0.27 variants when considering the V3-V4, V4, and V4-V5
regions of the gene, respectively. Although a species tended to have a
consistent number of 16S rRNA gene copies per genome, the number of
total variants increased with the number of genomes that were sampled
(Figure S1). For example, the 271 genome accessions of
\emph{Mycobacterium tuberculosis} in the \emph{rrn}DB each had 1 copy of
the gene per genome. However, across those accessions, there were 17
versions of the gene. An \emph{E. coli} genome typically had 7 copies of
the 16S rRNA gene with a median of 5 distinct full length ASVs per
genome (intraquartile range between 3 and 6). Across the 1,390 \emph{E.
coli} genomes in the \emph{rrn}DB, there were 1,402 versions of the
gene. These observations highlight the risk of selecting a threshold for
defining clusters that is too narrow because it is possible to split a
single genome into multiple clusters.

A method to avoid splitting a single genome into multiple clusters is to
cluster 16S rRNA gene sequences together based on their distances
between each other. Therefore, I assessed the impact of the distance
threshold used to define clusters of 16S rRNA genes on the propensity to
split a genome into separate clusters. To control for uneven
representation of genomes across species, I randomly selected one genome
from each species and repeated each randomization 100 times. I observed
that as the \emph{rrn} copy number increased, the distance threshold
required to reduce the ASVs in each genome to a single OTU increased
(Figure 1). Among species with 7 copies of the \emph{rrn} operon (e.g.,
\emph{E. coli}), a distance threshold of 5.25\% was required to reduce
full length ASVs into a single OTU for 95\% of the species. Similarly,
thresholds of 5.25, 2.50, and 3.75\% were required for the V3-V4, V4,
and V4-V5 regions, respectively. But, if a 3\% distance threshold was
used, then ASVs from genomes containing fewer than 6, 6, 8, and 6 copies
of the \emph{rrn} operon would reliably be clustered into a single OTU
for ASVs from the V1-V9, V3-V4, V4, and V4-V5 regions, respectively.
Consequently, these results demonstrate that broad thresholds must be
used to avoid splitting different operons from the same genome into
separate clusters.

At broad thresholds, 16S rRNA gene sequences from multiple species could
be clustered into the same ASV or OTU. I again randomly selected one
genome from each species to control for uneven representation of genomes
across species. For this analysis I measured the percentage of ASVs and
OTUs that contained 16S rRNA gene sequences from multiple species
(Figure 2). Without using distance-based thresholds, 4.1\% of the ASVs
contained sequences from multiple species when considering full length
sequences and 10.9, 16.2, and 13.1\% when considering the V3-V4, V4, and
V4-V5 regions, respectively. At the commonly used 3\% threshold for
defining OTUs, 27.4\% of the OTUs contained 16S rRNA gene sequences from
multiple species when considering full length sequences and 31.7, 34.3,
and 34.8\% when considering the V3-V4, V4, and V4-V5 regions,
respectively. Although the actual fraction of ASVs and OTUs that contain
sequences from multiple species is dependent on the taxonomic
composition of the sequences in the \emph{rrn}DB, this analysis
highlights the tradeoffs of using distance-based thresholds.

The results of this analysis demonstrate that there is a significant
risk of splitting a single genome into multiple clusters if using ASVs
or too fine of a threshold to define OTUs. An ongoing problem for
amplicon-based studies is defining a meaningful taxonomic unit (13, 15,
16). Since there is no consensus for a biologicaly definition of a
bacterial species (17--19), microbiologists must accept that how
bacterial species are named is biased and that taxonomic rules are not
applied in a consistent manner (e.g., (19, 20)). This makes it
impossible to fit a distance threshold to define an OTU definition that
matches a set of species names (21). Furthermore, the 16S rRNA gene does
not evolve at the same rate across all bacterial lineages (15), which
limits the biological interpretation of a common OTU definition. A
distance-based definition of a taxonomic unit based on 16S rRNA gene or
full genome sequences is operational and not necessarily grounded in
biological theory (15, 22--24). One benefit of a distance-based OTU
definition is the ability to mask residual sequencing error. The
analysis in this study was conducted using ideal sequences from
assembled genomes whereas sequences generated in microbiome studies
would harbor PCR and sequencing errors. These errors would only
exacerbate the inflated number of ASVs. There is general agreement in
bacterial systematics that to classify an organism to a bacterial
species, phenotypic and genome sequence data are needed (17--20). A
short sequence from a bacterial genome simply cannot differentiate
between species. Moreover, it is difficult to defend a clustering
threshold that would split a single genome into multiple taxonomic
units. It is not biologically plausible to entertain the possibility
that different \emph{rrn} operons from the same genome would have
different ecologies. Individual bacteria are defined at the cellular or
chromosomal level and not at the gene level. One could argue that, in
practice, communities are compared on a relative rather than absolute
basis. However, communities harboring populations that tend to have more
copies of the \emph{rrn} operon would appear to have higher richness and
diversity than those with fewer copies purely due to the propensity for
populations with more \emph{rrn} operons to generate more ASVs. Although
there are multiple reasons that proponents favor ASVs, the significant
risk of artificially splitting genomes into separate clusters is too
high to warrant their use.

\vspace{10mm}

\textbf{Materials and Methods. (i) Data availability.} The 16S rRNA gene
sequences used in this study were obtained from the \emph{rrn}DB
(\url{https://rrndb.umms.med.umich.edu}; version 5.7, released January
18, 2021) (14). At the time of submission, this was the most current
version of the database. The \emph{rrn}DB obtained the curated 16S rRNA
gene sequences from the KEGG database, which ultimately obtained them
from NCBI's non-redundant RefSeq database. The \emph{rrn}DB provided
downloadable versions of the sequences with their taxonomy as determined
using the naive Bayesian classifier trained on the RDP reference
taxonomy. For some genomes this resulted in multiple classifications
since a genome's 16S rRNA gene sequences were not identical. Instead, I
mapped the RefSeq accession number for each genome in the database to
obtain a single taxonomy for each genome. Because strain names were not
consistently given to genomes across bacterial species, I disregarded
the strain level designations.

\textbf{(ii) Definition of regions within the 16S rRNA gene.} The full
length 16S rRNA gene sequences were aligned to a SILVA reference
alignment of the 16S rRNA gene (v. 138) using the mothur software
package (v. 1.44.2) (25, 26). Regions of the 16S rRNA gene were selected
because of their use in the microbial ecology literature. Full length
sequences corresponded to \emph{E. coli} str. K-12 substr. MG1655
(NC\_000913) positions 28 through 1491, V4 to positions 534 through 786,
V3-V4 to positions 358 through 786, and V4-V5 to positions 534 through
908. The positions between these coordinates reflect the fragments that
would be amplified using commonly used PCR primers.

\textbf{(iii) Clustering sequences into OTUs.} Pairwise distances
between sequences were calculated using the dist.seqs command from
mothur. The OptiClust algorithm, as implemented in mothur, was used to
assign 16S rRNA gene sequences to OTUs (27). Distance thresholds between
0.25 and 10.00\% in 0.25 percentage point increments were used to assign
sequences to OTUs.

\textbf{(iv) Controlling for uneven sampling of genomes by species.}
Because of the uneven distribution of genome sequences across species I
randomly selected one genome from each species for the analysis of
splitting genomes and clustering ASVs from different species (Figures 1
and 2). The random selection was repeated 100 times. Analyses based on
this randomization reported the median of the 100 randomizations. The
intraquartile range between randomizations was less than 0.0024. Because
the range was so small, the confidence intervals were more narrow than
the thickness of the lines in Figures 1 and 2 and were not included.

\textbf{(v) Reproducible data analysis.} The code to perform the
analysis in this manuscript and its history are available as a git-based
version control repository on GitHub
(\url{https://github.com/SchlossLab/Schloss_rrnAnalysis_mSphere_2021}).
The analysis can be regenerated using a GNU Make-based workflow that
made use of built-in bash tools (v. 3.2.57), mothur (v. 1.44.2), and R
(v. 4.1.0). Within R, I used the tidyverse (v. 1.3.1), data.table (v.
1.14.0), Rcpp (v. 1.0.6), furrr (v. 0.2.2), here (v. 1.0.1) and
rmarkdown (v. 2.8) packages. The conception and development of this
analysis is available as a playlist on the Riffomonas YouTube channel
(\url{https://youtube.com/playlist?list=PLmNrK_nkqBpL7m_tyWdQgdyurerttCsPY}).

\textbf{(vi) Note on usage of ASV, OTU, and cluster.} I used ``ASV'' to
denote the cluster of true 16S rRNA gene sequences that were identical
to each other and ``OTU'' to denote the product of distance-based
clustering of sequences. Although ASVs do represent a type of
operational defition of a taxonomic unit and can be thought of as an OTU
formed using a distance of zero, proponents of the ASV approach prefer
to avoid the term OTU given the long history of OTUs being formed by
distance-based clustering
(\url{https://github.com/benjjneb/dada2/issues/62}; accessed
2021-02-26). For this reason, when an ASV split a genome into different
units, those units were called clusters rather than OTUs.

\vspace{10mm}

\textbf{Acknowledgements.} I am grateful to Robert Hein and Thomas
Schmidt, who maintain the \emph{rrn}DB, for their help in understanding
the curation of the database and for making the 16S rRNA gene sequences
and related metadata publicly available. I am also grateful to community
members who watched the serialized version of this analysis on YouTube
and provided suggestions and questions over the course of the
development of this project. This work was supported, in part, through
grants from the NIH (P30DK034933, U01AI124255, and R01CA215574).

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}

\noindent

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-Lane1985}{}%
1. Lane DJ, Pace B, Olsen GJ, Stahl DA, Sogin ML, Pace NR. 1985. Rapid
determination of 16S ribosomal RNA sequences for phylogenetic analyses.
\emph{Proceedings of the National Academy of Sciences} 82:6955--6959.
doi:
\href{https://doi.org/10.1073/pnas.82.20.6955}{10.1073/pnas.82.20.6955}.

\leavevmode\hypertarget{ref-Stackebrandt1994}{}%
2. Stackebrandt E, Goebel BM. 1994. Taxonomic note: A place for DNA-DNA
reassociation and 16S rRNA sequence analysis in the present species
definition in bacteriology. \emph{International Journal of Systematic
and Evolutionary Microbiology} 44:846--849. doi:
\href{https://doi.org/10.1099/00207713-44-4-846}{10.1099/00207713-44-4-846}.

\leavevmode\hypertarget{ref-Goris2007}{}%
3. Goris J, Konstantinidis KT, Klappenbach JA, Coenye T, Vandamme P,
Tiedje JM. 2007. DNA-DNA hybridization values and their relationship to
whole-genome sequence similarities. \emph{International Journal of
Systematic and Evolutionary Microbiology} 57:81--91. doi:
\href{https://doi.org/10.1099/ijs.0.64483-0}{10.1099/ijs.0.64483-0}.

\leavevmode\hypertarget{ref-RodriguezR2018}{}%
4. Rodriguez-R LM, Castro JC, Kyrpides NC, Cole JR, Tiedje JM,
Konstantinidis KT. 2018. How much do rRNA gene surveys underestimate
extant bacterial diversity? \emph{Applied and Environmental
Microbiology} 84:e00014--18. doi:
\href{https://doi.org/10.1128/aem.00014-18}{10.1128/aem.00014-18}.

\leavevmode\hypertarget{ref-Stackebrandt2006}{}%
5. Stackebrandt E, Ebers J. 2006. Taxonomic parameters revisited:
Tarnished gold standards. \emph{Microbiol Today} 33:152--155.

\leavevmode\hypertarget{ref-Edgar2018}{}%
6. Edgar RC. 2018. Updating the 97\% identity threshold for 16S
ribosomal RNA OTUs. \emph{Bioinformatics} 34:2371--2375. doi:
\href{https://doi.org/10.1093/bioinformatics/bty113}{10.1093/bioinformatics/bty113}.

\leavevmode\hypertarget{ref-Edgar2016}{}%
7. Edgar RC. 2016. UNOISE2: Improved error-correction for illumina 16S
and its amplicon sequencing. \emph{bioRxiv}. doi:
\href{https://doi.org/10.1101/081257}{10.1101/081257}.

\leavevmode\hypertarget{ref-Amir2017}{}%
8. Amir A, McDonald D, Navas-Molina JA, Kopylova E, Morton JT, Zech Xu
Z, Kightley EP, Thompson LR, Hyde ER, Gonzalez A, Knight R. 2017. Deblur
rapidly resolves single-nucleotide community sequence patterns.
\emph{mSystems} 2:e00191--16. doi:
\href{https://doi.org/10.1128/mSystems.00191-16}{10.1128/mSystems.00191-16}.

\leavevmode\hypertarget{ref-Callahan2016}{}%
9. Callahan BJ, McMurdie PJ, Rosen MJ, Han AW, Johnson AJA, Holmes SP.
2016. DADA2: High-resolution sample inference from illumina amplicon
data. \emph{Nature Methods} 13:581--583. doi:
\href{https://doi.org/10.1038/nmeth.3869}{10.1038/nmeth.3869}.

\leavevmode\hypertarget{ref-Eren2014}{}%
10. Eren AM, Morrison HG, Lescault PJ, Reveillaud J, Vineis JH, Sogin
ML. 2014. Minimum entropy decomposition: Unsupervised oligotyping for
sensitive partitioning of high-throughput marker gene sequences.
\emph{The ISME Journal} 9:968--979. doi:
\href{https://doi.org/10.1038/ismej.2014.195}{10.1038/ismej.2014.195}.

\leavevmode\hypertarget{ref-Pei2010}{}%
11. Pei AY, Oberdorf WE, Nossa CW, Agarwal A, Chokshi P, Gerz EA, Jin Z,
Lee P, Yang L, Poles M, Brown SM, Sotero S, DeSantis T, Brodie E, Nelson
K, Pei Z. 2010. Diversity of 16S rRNA genes within individual
prokaryotic genomes. \emph{Applied and Environmental Microbiology}
76:3886--3897. doi:
\href{https://doi.org/10.1128/aem.02953-09}{10.1128/aem.02953-09}.

\leavevmode\hypertarget{ref-Sun2013}{}%
12. Sun D-L, Jiang X, Wu QL, Zhou N-Y. 2013. Intragenomic heterogeneity
of 16S rRNA genes causes overestimation of prokaryotic diversity.
\emph{Applied and Environmental Microbiology} 79:5962--5969. doi:
\href{https://doi.org/10.1128/aem.01282-13}{10.1128/aem.01282-13}.

\leavevmode\hypertarget{ref-Callahan2017}{}%
13. Callahan BJ, McMurdie PJ, Holmes SP. 2017. Exact sequence variants
should replace operational taxonomic units in marker-gene data analysis.
\emph{The ISME Journal} 11:2639--2643. doi:
\href{https://doi.org/10.1038/ismej.2017.119}{10.1038/ismej.2017.119}.

\leavevmode\hypertarget{ref-Stoddard2014}{}%
14. Stoddard SF, Smith BJ, Hein R, Roller BRK, Schmidt TM. 2014.
\emph{rrn}DB: Improved tools for interpreting rRNA gene abundance in
bacteria and archaea and a new foundation for future development.
\emph{Nucleic Acids Research} 43:D593--D598. doi:
\href{https://doi.org/10.1093/nar/gku1201}{10.1093/nar/gku1201}.

\leavevmode\hypertarget{ref-Schloss2011}{}%
15. Schloss PD, Westcott SL. 2011. Assessing and improving methods used
in operational taxonomic unit-based approaches for 16S rRNA gene
sequence analysis. \emph{Applied and Environmental Microbiology}
77:3219--3226. doi:
\href{https://doi.org/10.1128/aem.02810-10}{10.1128/aem.02810-10}.

\leavevmode\hypertarget{ref-Johnson2019}{}%
16. Johnson JS, Spakowicz DJ, Hong B-Y, Petersen LM, Demkowicz P, Chen
L, Leopold SR, Hanson BM, Agresta HO, Gerstein M, Sodergren E, Weinstock
GM. 2019. Evaluation of 16S rRNA gene sequencing for species and
strain-level microbiome analysis. \emph{Nature Communications} 10:5029.
doi:
\href{https://doi.org/10.1038/s41467-019-13036-1}{10.1038/s41467-019-13036-1}.

\leavevmode\hypertarget{ref-Staley2006}{}%
17. Staley JT. 2006. The bacterial species dilemma and the
genomicphylogenetic species concept. \emph{Philosophical Transactions of
the Royal Society B: Biological Sciences} 361:1899--1909. doi:
\href{https://doi.org/10.1098/rstb.2006.1914}{10.1098/rstb.2006.1914}.

\leavevmode\hypertarget{ref-Oren2013}{}%
18. Oren A, Garrity GM. 2013. Then and now: A systematic review of the
systematics of prokaryotes in the last 80~years. \emph{Antonie van
Leeuwenhoek} 106:43--56. doi:
\href{https://doi.org/10.1007/s10482-013-0084-1}{10.1007/s10482-013-0084-1}.

\leavevmode\hypertarget{ref-Sanford2021}{}%
19. Sanford RA, Lloyd KG, Konstantinidis KT, Löffler FE. 2021. Microbial
taxonomy run amok. \emph{Trends in Microbiology}. doi:
\href{https://doi.org/10.1016/j.tim.2020.12.010}{10.1016/j.tim.2020.12.010}.

\leavevmode\hypertarget{ref-Baltrus2016}{}%
20. Baltrus DA, McCann HC, Guttman DS. 2016. Evolution, genomics and
epidemiology of \emph{Pseudomonas syringae}. \emph{Molecular Plant
Pathology} 18:152--168. doi:
\href{https://doi.org/10.1111/mpp.12506}{10.1111/mpp.12506}.

\leavevmode\hypertarget{ref-Konstantinidis2005}{}%
21. Konstantinidis KT, Tiedje JM. 2005. Towards a genome-based taxonomy
for prokaryotes. \emph{Journal of Bacteriology} 187:6258--6264. doi:
\href{https://doi.org/10.1128/jb.187.18.6258-6264.2005}{10.1128/jb.187.18.6258-6264.2005}.

\leavevmode\hypertarget{ref-Barco2020}{}%
22. Barco RA, Garrity GM, Scott JJ, Amend JP, Nealson KH, Emerson D.
2020. A genus definition for bacteria and archaea based on a standard
genome relatedness index. \emph{mBio} 11:02475--19. doi:
\href{https://doi.org/10.1128/mbio.02475-19}{10.1128/mbio.02475-19}.

\leavevmode\hypertarget{ref-Parks2020}{}%
23. Parks DH, Chuvochina M, Chaumeil P-A, Rinke C, Mussig AJ, Hugenholtz
P. 2020. A complete domain-to-species taxonomy for bacteria and archaea.
\emph{Nature Biotechnology} 38:1079--1086. doi:
\href{https://doi.org/10.1038/s41587-020-0501-8}{10.1038/s41587-020-0501-8}.

\leavevmode\hypertarget{ref-Yarza2014}{}%
24. Yarza P, Yilmaz P, Pruesse E, Glöckner FO, Ludwig W, Schleifer K-H,
Whitman WB, Euzéby J, Amann R, Rosselló-Móra R. 2014. Uniting the
classification of cultured and uncultured bacteria and archaea using 16S
rRNA gene sequences. \emph{Nature Reviews Microbiology} 12:635--645.
doi: \href{https://doi.org/10.1038/nrmicro3330}{10.1038/nrmicro3330}.

\leavevmode\hypertarget{ref-Schloss2009}{}%
25. Schloss PD, Westcott SL, Ryabin T, Hall JR, Hartmann M, Hollister
EB, Lesniewski RA, Oakley BB, Parks DH, Robinson CJ, Sahl JW, Stres B,
Thallinger GG, Horn DJV, Weber CF. 2009. Introducing mothur:
Open-source, platform-independent, community-supported software for
describing and comparing microbial communities. \emph{Applied and
Environmental Microbiology} 75:7537--7541. doi:
\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

\leavevmode\hypertarget{ref-Quast2012}{}%
26. Quast C, Pruesse E, Yilmaz P, Gerken J, Schweer T, Yarza P, Peplies
J, Glöckner FO. 2012. The SILVA ribosomal RNA gene database project:
Improved data processing and web-based tools. \emph{Nucleic Acids
Research} 41:D590--D596. doi:
\href{https://doi.org/10.1093/nar/gks1219}{10.1093/nar/gks1219}.

\leavevmode\hypertarget{ref-Westcott2017}{}%
27. Westcott SL, Schloss PD. 2017. OptiClust, an improved method for
assigning amplicon-based sequence data to operational taxonomic units.
\emph{mSphere} 2:00073--17. doi:
\href{https://doi.org/10.1128/mspheredirect.00073-17}{10.1128/mspheredirect.00073-17}.
\end{cslreferences}

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

\newpage

\textbf{Figure 1. The distance threshold required to prevent the
splitting of genomes into multiple OTUs increased as the number of
\emph{rrn} operons in the genome increased.} Each line represents the
median distance threshold for each region of the 16S rRNA gene that is
required for 95\% of the genomes with the indicated numbrer of
\emph{rrn} operons to cluster their ASVs to a single OTU. The median
distance threshold was calculated across 100 randomizations in which one
genome was sampled from each species. Only those number of \emph{rrn}
operons that were found in more than 100 species are included.

\textbf{Figure 2. As the distance threshold used to define an OTU
increased, the percentage of ASVs and OTUs representing multiple species
increased.} These data represent the median fractions for both
measurements across 100 randomizations. In each randomization, one
genome was sampled from each species.

\textbf{Figure S1. The ratio of number of distinct ASVs per copy of the
\emph{rrn} operon increased for a species as the number of genomes in
the \emph{rrn}DB for that species increased.} Each point represents a
different species and was shaded to be 80\% transparent so that when
points overlap they become darker. The blue line represents a smoothed
fit through the data. Both axes use a logarithmic scale (base 10).

\end{document}